If square dancing is the peak of order, geometry, and external command, its perfect chaotic counterpart would be Moshing (or the "Mosh Pit").

While square dancing relies on a "caller" to synchronize everyone into rigid patterns, moshing relies on individual impulse, physical impact, and organic flow.

In a square dance, you are a component of a machine. If you fail to follow the caller, the machine stops. In a mosh pit, you are a molecule in a gas. You collide, bounce, and redirect based on the energy of the crowd. It is the ultimate expression of entropy in dance.

If you first hear of a line, then you might think of a Euclidiean line.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 5
#| fig-cap: "foo"
#| label: fig-euclidiean-line

ggplot() +
  theme_void() +
  ## coord_equal() +
  ## theme_minimal() +
  scale_x_continuous(breaks = seq(-5, 5, 1), limits = c(0.5, 4.5)) +
  scale_y_continuous(breaks = seq(-5, 5, 1), limits = c(0.5, 4.5)) +
  #annotate("segment", x = 0.25, y = 0.5, xend = 5, yend = 0.5) +
  #annotate("segment", x = 0.5, y = 0.25, xend = 0.5, yend = 5) +  
  annotate("segment", x = 1, y = 1.5, xend = 4, yend = 4) +
  annotate("point", x = c(1, 4), y = c(1.5, 4)) +
  annotate("text", x = c(1, 4), y = c(1.5, 4) - 0.5, label = c("A", "B")) +
  labs(title = "This is a Euclidean line")

```


A possible line through points. Is this the best possible line?
  
  
```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 5
#| fig-cap: "foo"
#| label: fig-euclidiean-line-01

ggplot() +
  theme_void() +
  coord_equal() +
  theme_minimal() +
  scale_x_continuous(breaks = seq(-5, 5, 1), limits = c(0, 5)) +
  scale_y_continuous(breaks = seq(-5, 5, 1), limits = c(0, 5)) +
  annotate("segment", x = 0.25, y = 0.5, xend = 5, yend = 0.5,
           arrow = arrow(length = unit(0.02, "npc"), type = "closed")) +
  annotate("segment", x = c(1.5, 2.5, 3.5, 4.5), xend = c(1.5, 2.5, 3.5, 4.5),
           y = 0.45, yend = 0.55) +
  annotate("segment", x = 0.5, y = 0.25, xend = 0.5, yend = 5,
           arrow = arrow(length = unit(0.02, "npc"), type = "closed")) +
  annotate("segment", y = c(1.5, 2.5, 3.5, 4.5), yend = c(1.5, 2.5, 3.5, 4.5),
           x = 0.45, xend = 0.55) +
  ## line
  annotate("segment", x = 0.5, y = 2, xend = 5, yend = 4.5) +
  ## points
  annotate("point", x = c(1.5, 2, 3, 4.5), y = c(2, 3.5, 4, 3.5)) +
  ## parameters
  annotate("point", x = 0.5, y = 2, color = "red", size = 3) +
  annotate("segment", x = 2.5, y = 2+1.111, xend = 3.5, yend = 2+1.111,
           color = "red", size = 1) +
  annotate("segment", x = 3.5, y = 2+1.111, xend = 3.5, yend = 2+1.666,
           color = "red", size = 1) +
  annotate("segment", x = c(1.5, 2, 3, 4.5), y = c(2, 3.5, 4, 3.5), 
           xend = c(1.5, 2, 3, 4.5), yend = 2 + (c(1.5, 2, 3, 4.5)-0.5) * 0.55,
           color = "red", size = 1, linetype = "11") +
  ## box
  geom_shape(data = tibble(x = c(3, 4.75, 4.75, 3), 
                           y = c(0.75, 0.75, 2.75, 2.75)),
             aes(x, y), radius = unit(0.2, 'cm'), fill = "gray95") +
  annotate("text", x = (3+4.75)/2, y = 2.5, label = "Parameters",
           fontface = 2) +
  annotate("text", x = 3.25, y = c(2, 1.5, 1), 
           label = c("1 Intercept", "1 Slope", "n Deviations"),
           fontface = 3, hjust = "left") +
  annotate("text", x = 4.5, y = c(2, 1.5, 1), 
           label = c(expression(beta[0]), expression(beta[1]), expression(epsilon)),
           fontface = 3, hjust = "left") +
  labs(title = "This is a linear line")

```