Here comes the preface text

If square dancing is the peak of order, geometry, and external command, its perfect chaotic counterpart would be Moshing (or the "Mosh Pit").

While square dancing relies on a "caller" to synchronize everyone into rigid patterns, moshing relies on individual impulse, physical impact, and organic flow.

In a square dance, you are a component of a machine. If you fail to follow the caller, the machine stops. In a mosh pit, you are a molecule in a gas. You collide, bounce, and redirect based on the energy of the crowd. It is the ultimate expression of entropy in dance.

If you first hear of a line, then you might think of a Euclidiean line.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 5
#| fig-cap: "Visualisation of a line through two points, also known as a Euclidean line. Most people think of such a line first when they think about lines through points. This is mathematically correct, but not in statistics."
#| label: fig-euclidiean-line

ggplot() +
  theme_void() +
  ##coord_equal() +
  ##theme_minimal() +
  scale_x_continuous(breaks = seq(-5, 5, 1), limits = c(-0.25, 5)) +
  scale_y_continuous(breaks = seq(-5, 5, 1), limits = c(-0.25, 2.25)) +
  annotate("segment", x = -0.15, y = 0, xend = 5, yend = 0,
           arrow = arrow(length = unit(0.04, "npc"), type = "closed")) +
  annotate("segment", x = c(1, 2, 3, 4), xend = c(1, 2, 3, 4),
           y = -0.05, yend = 0.05) +
  annotate("segment", x = 0, y = -0.2, xend = 0, yend = 2.25,
           arrow = arrow(length = unit(0.04, "npc"), type = "closed")) +  
  annotate("segment", x = -0.05, xend = 0.05,
           y = c(1, 2), yend = c(1, 2)) +
  ## line
  annotate("segment", x = 0.5, y = 0.5, xend = 4.5, yend = 1.5, color = "gray50",
           size = 1) +
  annotate("point", x = c(0.5, 4.5), y = c(0.5, 1.5)) +
  annotate("text", x = c(0.5, 4.5), y = c(0.5, 1.5) - 0.25, label = c("A", "B"),
           size = 5, fontface = 2) +
  labs(title = "This is a Euclidean line") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

```


A possible line through points. Is this the best possible line?


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3
#| fig-width: 5
#| fig-cap: "Visualisation of a linear line through many points is a key concept in statistics. The line rarely hits all the points, so there is always a deviation between the line and the points. The line is described by two coefficients, or parameters: the intercept, where the line passes the y-axis, and the slope, which is the rate of change of the line. We observe one intercept, one slope and four deviations."
#| label: fig-euclidiean-line-01

ggplot() +
  theme_void() +
  ##coord_equal() +
  ##theme_minimal() +
  scale_x_continuous(breaks = seq(-5, 5, 1), limits = c(0.25, 5)) +
  scale_y_continuous(breaks = seq(-5, 5, 1), limits = c(0.25, 5)) +
  annotate("segment", x = 0.4, y = 0.5, xend = 5, yend = 0.5,
           arrow = arrow(length = unit(0.02, "npc"), type = "closed")) +
  annotate("segment", x = c(1.5, 2.5, 3.5, 4.5), xend = c(1.5, 2.5, 3.5, 4.5),
           y = 0.45, yend = 0.55) +
  annotate("segment", x = 0.5, y = 0.3, xend = 0.5, yend = 5,
           arrow = arrow(length = unit(0.02, "npc"), type = "closed")) +
  annotate("segment", y = c(1.5, 2.5, 3.5, 4.5), yend = c(1.5, 2.5, 3.5, 4.5),
           x = 0.475, xend = 0.525) +
  ## line
  annotate("segment", x = c(1.5, 2, 3, 4.5), y = c(2, 3.5, 4, 3.5), 
           xend = c(1.5, 2, 3, 4.5), yend = 2 + (c(1.5, 2, 3, 4.5)-0.5) * 0.55,
           color = "#FCA636FF", size = 1, lineend = "round") +
  annotate("segment", x = 0.5, y = 2, xend = 5, yend = 4.5, color = "gray50",
           size = 1) +
  ## parameters
  annotate("point", x = 0.5, y = 2, fill = "#0D0887FF", size = 3, shape = 21) +
  annotate("segment", x = 2.25, y = 2+(2.25-0.5)*0.55, xend = 3.25, yend = 2+(2.25-0.5)*0.55,
           color = "#B12A9080", size = 1, linetype = "11") +
  annotate("segment", x = 3.25, y = 2+(2.25-0.5)*0.55, xend = 3.25, yend = 2+(3.25-0.5)*0.55,
           color = "#B12A9080", size = 1, linetype = "11") +
  annotate("segment", x = 2.25, y = 2+(2.25-0.5)*0.55, xend = 3.25, yend = 2+(3.25-0.5)*0.55,
           color = "#B12A90FF", size = 1) +
  annotate("point", x = 2.75, y = 2+(2.75-0.5)*0.55, fill = "#B12A90FF", size = 3, shape = 21) +  
  ## points
  annotate("point", x = c(1.5, 2, 3, 4.5), y = c(2, 3.5, 4, 3.5)) +
  ## box
  geom_shape(data = tibble(x = c(3, 4.75, 4.75, 3), 
                           y = c(0.75, 0.75, 2.75, 2.75)),
             aes(x, y), radius = unit(0.2, 'cm'), fill = "gray95") +
  annotate("text", x = (3+4.75)/2, y = 2.5, label = "Parameters",
           fontface = 2, size = 5) +
  annotate("text", x = 3.1, y = c(2, 1.5, 1), size = 4.5,
           label = c("1 Intercept", "1 Slope", "n Deviations"),
           fontface = 4, hjust = "left", color = c("#0D0887FF", "#B12A90FF", "#FCA636FF")) +
  annotate("text", x = 4.5, y = c(2, 1.5, 1), 
           label = c(expression(beta[0]), expression(beta[1]), expression(epsilon)),
           size = 4.5, hjust = "center") +
  labs(title = "This is a linear line")  +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

```