```{r echo = FALSE}
pacman::p_load(tidyverse, ggforce)
```

# What is science? {#sec-what-is-science}

*Last modified on `r format(fs::file_info("chapter-01-what-is-science.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Reality is negotiable." --- Tim Ferriss*



```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-vulcan
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 7
#| fig-cap: "foo*"



ggplot() +
  #theme_void() +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 10.5, b = 4.9, angle = pi/5.5),
               color = "gray50", fill = "blue") +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 10, b = 5, angle = pi/5.5),
               color = "gray50", fill = "white") +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 5.5, b = 3.5, angle = -2.25),
               color = "gray50", linetype = 11) +
  coord_cartesian(xlim = c(-10, 10), ylim = c(-10,10)) +
  annotate("point", x = 0, y = 0, size = 10) +
  annotate("text", x = 0.5, y = 1, hjust = "left", label = "Sun", size = 4) +
  annotate("point", x = 8.6, y = 5, size = 3) +
  annotate("point", x = 3.5, y = 4.2, size = 3) +
  annotate("text", x = 3.7, y = 4.7, hjust = "left", label = "Vulcan", size = 4) 


```



## General background


> *"Any sufficiently advanced technology is indistinguishable from magic." ---  @clarke1968clarke, Clarke's third law*

Full quote

> *"Reality is negotiable. Outside of science and law, all rules can be bent or broken, and it doesn’t require being unethical." --- Tim Ferriss*

Idea of hypotheses

Science is guessing and falsification

@chalmers2013thing [What is this thing called Science?](https://www.pmf.unizg.hr/_download/repository/1._Dodatna_literatura_%28filozofija_znanosti%29%5B3%5D.pdf)

@feynman1966science [What is science](https://www-oc.chemie.uni-regensburg.de/diaz/media/docs/feynman_what-is-science_68.pdf)

@campbell1952science [What is science?](https://www.whatisscience.info/files/what_is_science_rd1.pdf)

@clark2025models [Models Demystified: A Practical Guide from Linear Regression to Deep Learning](https://m-clark.github.io/book-of-models/)

@cox2017statistical [Statistical Thinking for the 21st Century](https://statsthinking21.github.io/statsthinking21-core-site/index.html)

@deutsch2011beginning [The beginning of infinity: Explanations that transform the world](https://www.academia.edu/download/106176299/rtx120400544p.pdf)

[David Deutsch > Quotes](https://www.goodreads.com/author/quotes/103190.David_Deutsch)

## Theoretical background

## R packages used

## Data

## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Data Science" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}