```{r echo = FALSE, warning = FALSE, message = FALSE}
source("init.R")
source("images/part_1/part_1_data.R")
```

# Blood meal of fleas {#sec-temp}

*Last modified on `r format(fs::file_info("chapter-11-covariate-data.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

## General background

## Theoretical background

## R packages used

## Data

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.25
#| fig-width: 5
#| fig-cap: "foo."
#| label: fig-science-myth

ggplot() +
  theme_void() +
  ##theme_minimal() +
  coord_cartesian(xlim = c(0.9, 9.1), ylim = c(0.9, 9.25)) +
  scale_x_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  geom_shape(data = tibble(x = c(1, 8, 9, 2), 
                           y = c(3, 3, 5, 5)),
             aes(x, y), fill = "gray95", color = "black", size = 1) +
  geom_shape(data = tibble(x = c(2, 9, 9, 2), 
                           y = c(5, 5, 9, 9)),
             aes(x, y), fill = "gray95", color = "black", size = 1) +
  geom_shape(data = tibble(x = c(1.5, 7.5, 7.5, 1.5), 
                           y = c(1.2, 1.2, 2.8, 2.8)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "#B12A9080", color =  "#B12A90FF",
             alpha = 0.1) +
  annotate("text", x = 1.75, y = 2.3, label = "Todays play:", size = 6, fontface = 3,
           color = "#B12A90FF", hjust = "left") +  
  annotate("text", x = 4.3, y = 2, label = "Science", size = 11, fontface = 4,
           color = "#B12A90FF", hjust = "left") +  
  annotate("segment", x = 1, y = 1, xend = 1, yend = 3, size = 1) +
  annotate("segment", x = 8, y = 1, xend = 8, yend = 3, size = 1) +
  annotate("segment", x = 9, y = 3, xend = 9, yend = 5, size = 1) +
  annotate("text", x = 5.5, y = 7, label = "MYTH", size = 38, fontface = 2,
           color = "white") +
  annotate("text", x = 5.5, y = 7, label = "MYTH", size = 38, fontface = 2,
           color = "gray75") +
  ## bubble
  geom_ellipse(aes(x0 = 5.5, y0 = 7.5, a = 1.5, b = 1, angle = 0), color = "#6A00A8FF",
               fill = "white", alpha = 0.5) +
  annotate("text", x = 5.5, y = 7.5, label = "Metaphor", size = 9, fontface = 4,
           color = "#6A00A8FF") +
  annotate("point", x = c(3.1, 3.25, 3.5, 3.8), y = c(6.4, 6.65, 7, 7.25), size = c(2, 3, 4, 5), color = "#6A00A8FF",
               fill = "white", alpha = 1, shape = 21) +
  annotate("point", x = c(7.4, 7.3, 7.1), y = c(6.35, 6.65, 7), size = c(2, 3, 4), color = "#6A00A8FF",
               fill = "white", alpha = 1, shape = 21) +
  ## people
  annotate("point", x = c(3, 7.5), y = c(5.9, 5.8), size = 9, color = "black") +
  annotate("segment", x = c(3, 7.5), xend = c(3, 7.5), y = c(6, 6), yend = c(4.5, 4.5),
           size = 3, lineend = "round") +
  ## legs
  annotate("segment", x = c(3, 3), xend = c(2.8, 3.2), y = c(4.5, 4.5), yend = c(3.8, 3.8),
           size = 3, lineend = "round") +
  annotate("segment", x = c(7.5, 7.5), xend = c(7.3, 7.7), y = c(4.5, 4.5), yend = c(3.8, 3.8),
           size = 3, lineend = "round") +
  ## arms
  annotate("segment", x = c(3, 3), xend = c(2.75, 3.3), y = c(5.25, 5.25), yend = c(4.8, 5.5),
           size = 3, lineend = "round") +
  annotate("segment", x = c(7.5, 7.5), xend = c(7.25, 7.75), y = c(5.25, 5.25), yend = c(5.2, 4.8),
           size = 3, lineend = "round") 
  

```


### Linear

```{r}
jump_weight_tbl <- tibble(x = abs(rnorm(21, 3, 4)),
                          y = 10 + 1.2 * x + rnorm(21, 0, 2)) |>
  mutate_all(round, 1) |> 
  rename(weight = x, jumplength = y)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-01
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"

jump_weight_tbl |> 
  ggplot(aes(x = weight, y = jumplength, fill = col_pal(1, 0.5))) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21) +
  labs(x = "Body weight [mg]", y = "Jump length in [cm]",
       fill = "Host") +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  scale_x_continuous(breaks = 0:10) +
  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
```

```{r}
library(mvtnorm)

sigma <- matrix(c(4,2,2,3), ncol=2)
x <- rmvnorm(n = 21, mean=c(1,2), sigma=sigma) |> 
  as_tibble() |> 
  rename(weight = V1, bodylength = V2)
colMeans(x)
var(x)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-02
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"

x |> 
  ggplot(aes(x = weight, y = bodylength, fill = col_pal(1, 0.5))) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21) +
  labs(x = "Body weight [mg]", y = "Body length [mm]",
       fill = "Host") +
  scale_fill_viridis(discrete = TRUE, option = "plasma") #+
  # scale_x_continuous(breaks = 0:10) +
  # coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
```


### Non-linear


```{r}
jump_weight_non_linear_tbl <- tibble(x = abs(rnorm(32, 3, 4)),
                                     y = 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2 + rnorm(32, 0, 1)) |>
  rename(weight = x, jumplength = y)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-04
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"

jump_weight_non_linear_tbl |> 
  ggplot(aes(x = weight, y = jumplength, fill = col_pal(1, 0.5))) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21) +
  labs(x = "Body weight [mg]", y = "Jump length in [cm]",
       fill = "Host") +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  scale_x_continuous(breaks = 0:10) +
  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
```


## Data availability

The data is available as txt-Files under <https://github.com/jkruppa/biodatascience>.

```{r}
#| echo: false

jump_weight_tbl |> 
  mutate_if(is.numeric, round, 2) |> 
  write_delim("data/jump_weight.txt")

jump_weight_non_linear_tbl |> 
  mutate_if(is.numeric, round, 2) |> 
  write_delim("data/jump_weight_non_linear.txt")
```


## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}