```{r echo = FALSE, warning = FALSE, message = FALSE}
source("init.R")
source("images/part_5/part_5_model.R")
```

# It started with a line {#sec-started-line}

*Last modified on `r format(fs::file_info("chapter-52-it-started-with-a-line.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"It started with a ~~kiss~~ line; [...] How could I resist?" --- It Started With A Kiss by Hot Chocolate*

In statistics, everything is about a line. This is also a lie. More precisely, in parametric statistics, everything is about a line. In a later chapter, we will also explore nonparametric statistical analysis, which is different but somewhat outdated. We have better tools on your shelves to solve problems that were previously solved using nonparametric approaches.


### The square and the root

> *"What you see, well, you might not know; You get the feelin', comin' after the glow" --- The joker & the thief by Wolfmother*

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-square-and-root

p1_square <- ggplot() +
  theme_void() +
  coord_equal() +
  scale_x_continuous(breaks = seq(-5, 5, 1), limits = c(-1.25, 1.25)) +
  scale_y_continuous(breaks = seq(-5, 5, 1), limits = c(-1.25, 1)) +
  geom_tile(aes(x = 0, y = 0, width = 2, height = 2),
            fill = "#FCA63680", color = "#FCA636FF", alpha = 0.2, linewidth = 0.75) +
  annotate("text", x = 0, y = 0, label = expression(a^2), 
           size = 7.5, color = "black") +
  annotate("text", x = 0, y = -1.15, label = expression("a"), color = "#FCA636FF", 
           size = 6, fontface = 2) +
  annotate("text", x = 1.15, y = 0, label = expression("a"), color = "#FCA636FF", 
           size = 6, fontface = 2, angle = -90) +
  labs(title = "This is a square") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

p2_square_root <- ggplot() +
  theme_void() +
  coord_equal() +
  scale_x_continuous(breaks = seq(-5, 5, 1), limits = c(-1.25, 1.25)) +
  scale_y_continuous(breaks = seq(-5, 5, 1), limits = c(-1.25, 1)) +
  geom_tile(aes(x = 0, y = 0, width = 2, height = 2), linetype = 2,
            fill = "#FCA63680", color = "#FCA63680", alpha = 0.1, linewidth = 0.75) +
  annotate("segment", x = 1, y = -1, xend = 1, yend = 1, linewidth = 1.75, color = "#FCA636FF") +
  annotate("text", x = 1.15, y = 0, label = expression(sqrt(phantom(aa))), size = 6.5, 
           color = "#FCA636FF", angle = -90) +
  labs(title = "This is a root") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

p1_square + p2_square_root +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```



```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-square-sum

p1_square_many <- ggplot() +
  theme_void() +
  ## theme_minimal() +
  coord_equal() +
  scale_x_continuous(breaks = seq(-5, 5, 1), limits = c(-1, 1)) +
  scale_y_continuous(breaks = seq(-5, 5, 1), limits = c(-1, 1)) +
  geom_tile(aes(x = c(-0.5, -0.5, 0.5, 0.5), y = c(-0.5, 0.5, -0.5, 0.5), 
                width = c(0.5, 1, 0.75, 0.45), 
                height = c(0.5, 1, 0.75, 0.45)), linetype = 1,
            fill = "#FCA63680", color = "#FCA636FF", alpha = 0.2, linewidth = 0.75) +
  annotate("text", x = c(-0.5, -0.5, 0.5, 0.5), y = c(-0.5, 0.5, -0.5, 0.5), 
           label = c(expression(a[1]^2), expression(a[2]^2), expression(a[3]^2), expression(a[4]^2)), 
           size = 6, color = "black") +
  labs(title = "This are many squares") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

p2_square_sum <- ggplot() +
  theme_void() +
  ##theme_minimal() +
  coord_equal() +
  scale_x_continuous(breaks = seq(-5, 5, 1), limits = c(-1, 1)) +
  scale_y_continuous(breaks = seq(-5, 5, 1), limits = c(-1, 1)) +
  geom_tile(aes(x = 0, y = 0, width = 1.39, height = 1.39),
            fill = "#FCA63680", color = "#FCA636FF", alpha = 0.2, linewidth = 0.75) +
  annotate("text", x = 0, y = 0, size = 7.5, color = "black", label = expression(sum()*a[i]^2),
           fontface = 2) +
  annotate("text", x = 0, y = -0.875, label = expression(sqrt(sum())), 
           size = 5, color = "#FCA636FF") +
  annotate("text", x = 0.875, y = 0, label = expression(sqrt(sum())),
           size = 5, color = "#FCA636FF", angle = -90) +
  labs(title = "This is the sum of squares") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

p1_square_many + p2_square_sum +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

### The line and the square

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "Visualisation of a line through two points, also known as a Euclidean line. Most people think of such a line first when they think about lines through points. This is mathematically correct, but not in statistics."
#| label: fig-euclidiean-line

p1_euclidean_line + p2_line_squared +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```




## The hunter and the rabbit

> *"Statistics is: When the hunter misses the rabbit once to the left and once to the right, on average the rabbit is dead." --- Mike Krüger, German comedian*

I first came across the joke on Tweedback, a tool for anonymous feedback in lectures. One student had written the joke, which intrigued me. At first, I could not understand why. Then I understood the deeper meaning. In statistics, it's all about lines. Or lines through points. A gunshot is a line. A mean is a line. Single shots deviate from the average shot, as do single observations from the mean. If we measure the distance from the mean to each observation, we will get positive and negative values. Ultimately, these values would sum to zero. I don't believe for a moment that the comedian truly understands the depth of the joke from a statistical perspective. 

In the following @fig-rabbit-shot, I have illustrated the example with some numbers. The rabbit is sitting at position 5m. The hunter shoots at $3m$, which is $-2m$ to the left. The second shot goes to the right, at $7m$, which is a deviation of $+2m$. We can now calculate the average of the two shots. This is equal to $5m$. The hunter hits. When we calculate the deviation, we get zero: $(-2) + (+2) = 0$. Therefore, the whole hunt of two shots has no deviation at all. This is wrong. Our model of reality tells us something completely different compared to what our hunter experienced. Therefore, we need a better model.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2.25
#| fig-width: 7
#| fig-cap: "Illustration of the joke about the hunter and the rabbit. If the hunter shoots past the rabbit on the left and on the right, then on average the hunter will hit the rabbit. The deviation is zero. The joke illustrates the problem of measuring a statistical deviation when positive and negative values can occur."
#| label: fig-rabbit-shot

p_rabbit_joke
```


### Normal distributed data

> *"Can't explain all the feelings that you're making me feel; I believe in a thing called ~~love~~ normality." --- I Believe in a Thing Called Love by The Darkness*

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}