```{r echo = FALSE, warning = FALSE}
source("init.R")
source("images/part_0/part_0_model.R")
```

# What is a model? {#sec-what-is-a-model}

*Last modified on `r format(fs::file_info("chapter-03-what-is-a-model.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

Imagine you are standing on a globe but still believe it is flat. You are standing on solid ground, with boiling rocks a few kilometers below you. A few kilometers above, you cannot breathe anymore. You are not moving, but the Earth is moving at speeds of up to 828,000 km/h, depending on the reference point. Your reality is a serious misinterpretation.


## General background

| Symbol | Name | Application | Description |
|:--------------:|:-------------------|:---------------|:-------------------|
| $y$ | **outcome**, response, endpoint, dependent variable |  | *The right-hand side (abbr. RHS) of the model. Describing the values measured in an experiment or study.* |
| $x$ | **influencer**, influential variable, risk factor, fixed effect, independent variable |  | *The left-hand side (abbr. LHS) of the model. Describing the influential variables in an experiment or study.* |
| $z$ | **random effect** |  | *A factor that provides a description of an grouping variable, which is not part of the controlled experimental setting.* |
| $x$ | **explanator**, explanatory variable | Explanation | *The influencer is used to describe or explain the outcome.* |
| $x$ | **predictor**, predictive variable | Prediction | *The influencer is used to predict the outcome.* |
| $x$ | **focal explanator**, **focal predictor**, focal variable | Main effect | *In a model with multiple influencers, the focal variable is the variable of primary interest.* |
| $c$ | **covariate**, covariable | Continuous $x$ | *The influencer is a numeric variable with continuous values.* |
| $f_A$ | **factor A**, factorial variable, categorical variable | Categorical $x$ | *The influencer is discrete, functioning as a grouping variable, such as an experimental group or a treatment.* |
| $A.1$ to $A.j$ | **levels**, groups, treatment groups | Factor $f_A$ | *The discrete groups included in one factor A.* |

: Table of terms used in the statistical modelling. The terms in bold are used here. Depending on the scientific background, the usage of these terms can vary widely. {#tbl-names-regression tbl-colwidths="$$20, 5, 30, 45$$"}

A sentence why we use $y$ and not $x$ for mean and other stuff.

@mccullagh2002statistical [What is a statistical model?](https://projecteuclid.org/journals/annals-of-statistics/volume-30/issue-5/What-is-a-statistical-model/10.1214/aos/1035844977.pdf)

@appleton1995we [What do we mean by a statistical model?](https://onlinelibrary.wiley.com/doi/abs/10.1002/sim.4780140209?casa_token=jLDogF46zS8AAAAA%3AKOKHMTgbj4gK8ayVXmT1GaXq7slv9i3VGqzA184yydwKUjsDIszepgOaeMQc_L5hRY-TpLZvgDDIdDea)

@hand2019purpose [What Is the Purpose of Statistical Modeling?](https://hdsr.mitpress.mit.edu/pub/9qsbf3hz/release/2?from=603&to=744)

@spanos2006statistical [Where do statistical models come from? Revisiting the problem of specification](https://arxiv.org/pdf/math.ST/0610849)

@gilchrist1984statistical [Statistical modelling](https://nzdr.ru/data/media/biblio/kolxoz/M/MV/MVsa/Gilchrist%20W.%20Statistical%20modelling%20with%20quantile%20functions%20(Chapman,%202000)(ISBN%201584881747)(317s)_MVsa_.pdf)

## Theoretical background

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-1
p_lhs_rhs
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-2
p_model_abstract
```


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-4
p_model_exp_simple_reg
```


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-5
p_simple_model
```
```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-6
p_mult_model
```


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-7
p_lhs_rhs_simple_r
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-8
p_lhs_rhs_mult_r
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2.5
#| fig-width: 9.5
#| fig-cap: "foo"
#| label: fig-model-in-R-9
p_regression_wording
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-10
p_lhs_rhs_detail
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 2.5
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-model-in-R-11
p_problem_x 

c("#0D0887FF", "#6A00A8FF", "#B12A90FF", "#E16462FF", "#FCA636FF" ,"#F0F921FF")
c("#0D088780", "#6A00A880", "#B12A9080", "#E1646280", "#FCA63680", "#F0F92180")
```

> *"Twistin', shake it, shake it, shake it, shake it, baby; Hey, we gonna Loop de Loop
Shake it out, baby; Hey, we gonna Loop de La" --- [The Blues Brothers (1980) -- Shake a Tail Feather](https://www.youtube.com/watch?v=qdbrIrFxas0)*

> *"It started with a ~~kiss~~ line; [...] How could I resist?" --- [Hot Chocolate -- It Started With A Kiss](https://www.youtube.com/watch?v=NosS1rpJ7tU)*

## R packages used

## Data

## Alternatives

Further tutorials and R packages on XXX

## Glossary

term

:   what does it mean.

## The meaning of "Models of Reality" in this chapter.

-   itemize with max. 5-6 words

## Summary

## References {.unnumbered}
