```{r echo = FALSE, warning = FALSE, message = FALSE}
source("init.R")
source("images/part_0/part_0_science.R")
```

# What is science? {#sec-what-is-science}

*Last modified on `r format(fs::file_info("chapter-01-what-is-science.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Reality is negotiable." --- Tim Ferriss*

The first chapter of this book is already the hardest. I started writing this chapter very early on, perhaps even before any of the others. Yet it took the longest to finish. The question is simple: What is science? The more I researched, the more complicated the answer became. Or the other way around. I followed the advice of Wolfgang Pauli. Pauli defined the scientific method as "taking up a subject repeatedly. Thinking about it. Setting it aside. Gathering new empirical material. And continuing this process for years if necessary. In this way, the conscious mind stimulates the unconscious mind. If anything, this is the only way to achieve results." Therefore, I tried to read and find answers to the question of what science might be. Perhaps science is, at its core, something personal. However, science also has objective aspects that I want to present here. Ultimately, decide for yourself what science truly is.

The following @fig-overview-sience-non-science shows the demarcation line between science and non-science. In the next sections, we will discuss each specific part, but I think it's important to start with a clear overview. Science is distinguished from non-science through experimentation and falsification. This is true in this book. There are many subtle definitions, but I will keep it simple here. If an experiment is not possible, then it is not science. We also need the ability to build a falsification framework. Therefore, we need hypotheses that can be rejected by experimentation. These two requirements are met by the natural sciences, including physics, chemistry, and biology, in order of their purity. Biology includes medicine and all other fields of knowledge gathered from living things. In addition to natural science, there is formal science, which can be seen as a tool of science. This is due to the lack of experimentation, falsification, or both. Therefore, we will not include mathematics, logic, or theoretical computer science as sciences. True non-science fields are intuition-based or subjective at their core. There are many of these fields, and we will discuss them shortly. The humanities are undoubtedly different from non-science, but this book is not about the humanities, so we will put them aside.        

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 10
#| fig-cap: "Visualization of the demarcation line of science. This line divides science from non-science through experimentation and the principle of falsification. The purity of science fields is taken from xkcd comic #435 ([https://xkcd.com/435](https://xkcd.com/435/)). According to this definition, natural science, including physics, chemistry, and biology, is separated from formal science, including mathematics. Formal science must be clearly distinguished from other intuition-based and subjective non-science, such as pseudoscience or anti-science. The humanities are indisputably different but are also considered non-science in this schematic."
#| label: fig-overview-sience-non-science

p_science_overview 
```

Perhaps we should start with the supposedly easy part. What is the opposite of science? In doing so, we will gain an understanding of the term "science" and what it means to us. This may be the deepest lesson for me: we cannot really grasp science. Science is connected to each person in a different emotional way. Science seems cold and rational, but humans are not. Therefore, our relationship with science is strange and complex.

## What is the opposite of science?

As I found out, there is no direct opposite of science. That's interesting. Sometimes, you can define something by knowing what it is not. You might think this is not so problematic. Take a few minutes to ask yourself: "What is the opposite of science?" Okay, good to have your focus back on the page. Maybe you thought science was cold and rational. I will share my personal story with science. For me, science was Star Trek: Enterprise. But I'm too young for Kirk and Spock. So, I was socialized with Captain Jean-Luc Picard and Star Trek: The Next Generation (TNG). That's where I first encountered Shakespeare, but that's another story. In TNG, science was determined by the exploration of new worlds. I would sit comfortably on the open bridge in front of the gigantic viewscreen. Commander Data was the embodiment of science. He was an android. He was rational and cold, devoid of emotion and humor. Let's flip these attributes.

If an android is science, then what is a human? Then is the human the opposite of science? Perhaps there is some truth to that. Humans have lived on this planet for thousands of years, but science as we know it began during the Enlightenment. So perhaps humans are the opposite of science. Human beings are full of emotions and feelings. Each one is very subjective and interested in their own development. They feel love and romance. Neither of these is very scientific. Humans also believe. Some believe in a more structured manner, like in big churches, while others believe in a more individual manner. However, believing in something has nothing to do with science. But humans do science. Therefore, science is part of humanity. As Joshua Schimel said, "Knowledge is a product of human hopes and fears. Our emotions are crucial to its development, and it cannot be truly understood as some bloodless, emotionless enterprise."@schimel2012writing Science is a human endeavor.  

According to Arthur C. Clarke's Third Law, "Any sufficiently advanced technology is indistinguishable from magic."@clarke1968clarke Is magic therefore the opposite of science? We cannot perform magic. We do not have the power to do so. However, some humans believe that we can perform magic. Who is "we"? The advanced part of humankind. We come into contact with our brothers and sisters who are at a different technological level, such as tribal cultures on isolated islands in Indonesia. To a tribe with a primitive culture, our technology seems like magic. Here, we can observe the cargo cult. A cargo cult is named after its appearance in World War II. The US Army faced the challenge of determining the best island-hopping route to mainland Japan. Since no one knew and the tides of war were constantly changing, the US Army Logistics Corps decided to drop cargo everywhere. This way, if US soldiers approached an island, there would already be plenty of supplies. 

What happened to the indigenous population? They believed that the gods had come to bring them cargo. They could not think of any other explanation for what they observed. They observed hardened silver birds dropping and landing on the islands. After a while, the gods stopped showing up. The war was over. But nobody told them. Therefore, the people of the islands copied the gods' behavior. They built the control room, and they made a radar station out of grass and wood. They even built plans out of grass. Nothing could work. They did everything perfectly, but they missed the point entirely. It was a cargo cult, not science. They copied someone else's routine and hoped to experience what the gods experienced when they did the same.

Do you follow a cargo cult in your daily routines? Do you do things that you have seen but don't understand? Do you believe that they will work even though you don't know why? This is totally human and from my point of view totally necessary to be a working community member. Some stuff must be done, even if I don not really understand the why. But what if we do it in science? Then we will do cargo cult science @feynman1998cargo. It is a danger of enormous proportions if done in the wrong place with the wrong intentions. Still, most intentions are good and are twisted by the circumstances. 

There are several examples of following the rules dogmatic and get a catastrophically  outcome. Later we will visit different models of the reality including the Fukushima nuclear accident. 

Here I want to conclude, that the opposite of science is hard to grasp. On the one hand we can argue, that the emotions and feelings of humans are the opposite of the cold and rational science. But again, the emotions are the driving force of humans to do science. Therefore, we do something very scientific, we led the part open for further research and look for things that are not science. 

## What is not science?



Mathematics and psychoanalysis have no experiments.

Mathematics and logic are more like scientific tools to do science.

Romantic as the answer to enlightment.

conspiracy theories @douglas2023conspiracy and @grimes2016viability


## Why do science?

To find a good explanation

To help people

To reduce suffering and pain

To make the plant earth more habitable 

To live longer

Because the human is lazy@price2021laziness




## How to do science?

"As a scientist, you are a professional writer," Joshua Schimel said. @schimel2012writing

Pauli defined the scientific method as "taking up a subject repeatedly. Thinking about it. Setting it aside. Gathering new empirical material. And continuing this process for years if necessary. In this way, the conscious mind stimulates the unconscious mind. If anything, this is the only way to achieve results." @fischer2022stunde. This is more or less about taking a chair and sitting in the middle of the room and getting bored as long as possible. Doing so sparks creativity.

Otto Kruse stated in his book, "many scientific topics can only be solved once all aspects have been explicitly formulated. The mind is not sufficiently prepared for this, as it can only ever focus on small sections at a time. Systematic thinking is only possible when writing, i.e. recording the results of one's thinking and relating them to other aspects."@kruse2007keine

The two paths of logic.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 10
#| fig-cap: "The difference between the two logical pathways: On the left is the inductive logic, which is a bottom-up approach that goes from specific to general. First, we observe and collect data, then we look for patterns in the data. Then, we draw a general conclusion about the pattern. After observing something a few times, we conclude that it is probably a rule. On the right is the deductive logic, a top-down approach from general to specific. We have a general theory and conduct an experiment to test it. Then, we draw a conclusion. From a more common perspective, the two logical pathways have things in common but differ in their train of thought."
#| label: fig-sience-inductive

p_sience_inductive 
```


Erst Mach and observational things.

Karl Pearson what is reality.

Quantum mechanics, we do not know what reality is.

The measurement paradox. When there is no agent wo is measuring, does atoms exist? What does measuring mean? A interaction between a large and a small atomic object. Where is the border between large and quantum objects?

## What is statistics?

A toolbox of mathematical models to do science.

In this book we will use Fisherian statistics, which is inductive logic reasoning. Or in other words the logic of inductive inference.

Fisher switched from a mathematical deductive reasoning to a inductive reasoning in science. This is remarkable for me, because he changed the way he does science.

Statistics by Fisher and quantum mechanics have in common that both do inductive probabilistic reasoning. Scientific models are deductive

Statistical models are inductive


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 4.5
#| fig-cap: "Inductive reasoning is a probabilistic process in statistics. First, a hypothesis is stated. Then, an experiment is conducted based on the hypothesis, and data is collected. This data is a random representation of observables. This is where the probabilistic process begins. Then, statistical modeling is applied to the experimental data to minimize noise. Finally, a conclusion is drawn. Sometimes the aim is to make a probabilistic decision or explanation, with or without prediction."
#| label: fig-statistic-inductive

p_statistic_inductive 
```

The inductive reasoning seems a little bit hard to follow. Therefore I made a small experiment, so achieve the formula for the genetic energy of a object. The formula is well known science Newton found them. The formula for the kinetic energy is $E_[kinetic] = \tfrac{1}{2}mv^2$. Therefore we take the mass of a object in kilogram and multiply it by the quadratic velocity. The velocity is defined by the following $v = \tfrac{d}{t}$. We divide the traveled distance by the needed time of the object. Now I want to find the dependency between the velocity and the kinetic energy. Therefore I build the following apparatus. A ball with a mass of $30g$ or $0.03kg$ lies on a height adjustable tower. The ball will travel without friction a ramp with a distance of $60cm$ or $0.6m$ to the ground. We will measure at different heights how long the ball will take to reach the ground. In each trail we can than calculate the velocity. We know the energy of the ball on the top. There we find the potential energy $E[potential] = mgh$. We multiply the mass of the ball with the earth acceleration of $9.81\tfrac{m}{s^2}$ and the height of the tower. Because we change the height, we also increase the energy of the ball.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.75
#| fig-width: 5.75
#| fig-cap: "The formula for kinetic energy can be derived through the following example experiment. A ball of fixed mass rolls down a fixed ramp from different heights. For each trial, the time it takes the ball to reach the bottom is measured with a stopwatch. The velocity can be calculated from the traveled distance and time. As the height increases, the ball will reach the ground in less time. Friction can be ignored. However, a measurement error will most likely occur in each trial."
#| label: fig-newton-example

p_newton_example
```

Let write everything together. We have the mass of the ball $m = 0.03kg$ and the length of the ramp $d = 2m$ as well as the earth acceleration $g = 9.81 m/s^2$. Now we put the tower at five heights and let the ball roll. Surprisingly fast runs the ball. Because we can not perfectly adjust the height and the manual time measurement is error prone we get noisy measures of the velocity. We observe data. And obsserved data has an error.

| Height \[m\] | Potential energy \[J\] | Time \[s\] | Velocity \[m/s\] |
|--------------|------------------|--------------|--------------|
| $0.1$ | $0.1 \cdot 0.03 \cdot 9.81 = 0.029$ | $1.58$ | $2/1.58 = 1.27$ |
| $0.2$ | $0.059$ | $0.95$ | $2.11$ |
| $0.3$ | $0.088$ | $0.85$ | $2.36$ |
| $0.4$ | $0.118$ | $0.65$ | $3.09$ |
| $0.5$ | $0.147$ | $0.72$ | $2.78$ |

: Results table from the experiment to determine the formula for kinetic energy. The height was changed five times. Using the acceleration due to gravity, the potential energy of the ball can be determined from its height and fixed weight. The ball's speed can be determined from the rolling time and the fixed ramp length. {#tbl-newton-data}

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "Visualization of the relationship between energy and velocity. The points represent calculated velocities from different heights. Due to measurement errors, the points do not fit perfectly on a line. The blue line is the formula derived from the data using a nonlinear model. The shaded area shows the uncertainty caused by measurement errors. The purple line shows the true dependency using the formula for kinetic energy."
#| label: fig-kenetic-energy

m <- 0.03
d <- 2
t_func <- \(h) d/sqrt(2*9.81*h)

newton_tbl <- tibble(h = c(0.1, 0.2, 0.3, 0.4, 0.5),
       t = t_func(h) + rnorm(5, 0, 0.1),
       v = d/t,
       E = m * 9.81 * h) 

pacman::p_load(nlraa)

nls_fit <- nls(E ~ b0 * I(v^b1), data = newton_tbl, 
           start = c(b0 = 0.1, b1 = 1))

nls_ci <- predict_nls(nls_fit, interval = "prediction")

cbind(newton_tbl, nls_ci) |> 
  ggplot(aes(v, E)) +
  theme_book() +
  geom_function(fun = \(x) 0.5 * 0.03 * x^2, color = "#B12A90FF",  
                size = 1) +
  geom_function(fun = \(x) 0 + 0.021 * x^(1.69), color = "#0D0887FF",
                size = 1) +
  scale_x_continuous(limits = c(0, 3.5), breaks = seq(0, 3.5, 0.5)) +
  scale_y_continuous(limits = c(-0.02, 0.25), breaks = seq(0, 0.25, 0.05)) +
  geom_point(size = 4) +
  labs(x = "Velocity [m/s]", y = "Energy [J]") +
  annotate("text", hjust = "left", x = 0.5, y = 0.2, size = 5,
           label = expression("E"[kinetic]~"="~frac(1, 2) %.% m %.% v^2), color = "#B12A90FF") +
  annotate("text", hjust = "left", x = 0.5, y = 0.15, size = 5,
           label = expression("E"[model]~"="~frac(1, 1.43) %.% m %.% v^1.69), 
           color = "#0D0887FF") +
  geom_ribbon(aes(ymin = Estimate - Est.Error, ymax = Estimate + Est.Error), 
              fill = "#0D088780", alpha = 0.2) +
  annotate("text", x = 2.5, y = 0.0325, label = "Uncertainty area",
           color = "#0D0887FF", fontface = 4, size = 5)
```

## What is a good explanation?

## What is real?


Sometimes we need a model to know what is real.

Fukusima @sevon2020analysis @sevon2023analyzing


## Age of enlightment

## General background

> *"Any sufficiently advanced technology is indistinguishable from magic." --- Arthur C. Clarke's Third Law @clarke1968clarke*

Full quote

> *"Reality is negotiable. Outside of science and law, all rules can be bent or broken, and it doesn’t require being unethical." --- Tim Ferriss*

Idea of hypotheses

Science is guessing and falsification

@chalmers2013thing [What is this thing called Science?](https://www.pmf.unizg.hr/_download/repository/1._Dodatna_literatura_%28filozofija_znanosti%29%5B3%5D.pdf)

@feynman1966science [What is science](https://www-oc.chemie.uni-regensburg.de/diaz/media/docs/feynman_what-is-science_68.pdf)

@campbell1952science [What is science?](https://www.whatisscience.info/files/what_is_science_rd1.pdf)

@clark2025models [Models Demystified: A Practical Guide from Linear Regression to Deep Learning](https://m-clark.github.io/book-of-models/)

@cox2017statistical [Statistical Thinking for the 21st Century](https://statsthinking21.github.io/statsthinking21-core-site/index.html)

@deutsch2011beginning [The beginning of infinity: Explanations that transform the world](https://www.academia.edu/download/106176299/rtx120400544p.pdf)

[David Deutsch \> Quotes](https://www.goodreads.com/author/quotes/103190.David_Deutsch)

## Theoretical background

## R packages used

## Data

## Alternatives

Further tutorials and R packages on XXX

## Glossary

term

:   what does it mean.

## The meaning of "Models of Reality" in this chapter.

-   itemize with max. 5-6 words

## Summary

## References {.unnumbered}
