```{r echo = FALSE, warning = FALSE, message = FALSE}
source("init.R")
source("images/part_1/part_1_data.R")
```

# Body weight of fleas {#sec-temp}

*Last modified on `r format(fs::file_info("chapter-11-covariate-data.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

## General background

## Theoretical background

## R packages used

## Data

### Linear

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-01-flea
#| fig-align: center
#| fig-height: 3
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"
p_measured_flea
```
Quality Rating (ACR): 1-Bad, 2-Poor, 3-Fair, 4-Good, 5-Excellent.


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-01-flea-measure
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 12
#| fig-cap: "foo **(A)** foo **(B)** foo **(C)** foo"

p_single_obs_flea <- ggplot() +
  theme_minimal() +
  theme_void() +
  coord_cartesian(xlim = c(-10, 10), ylim = c(-10, 10)) +
  geom_image(aes(x = c(-7, -6, 0, 5, 7, -5), y = c(8, -7, -5, 0, 6, 2), 
                 image = rep("images/flea_00.png", 6)), size = 0.3) +
  annotate("text", x = c(-7, -6, 0, 5, 7, -5),
           y = c(8, -7, -5, 0, 6, 2)-2.5, fontface = 2, size = 4.5, color = "gray50",
           label = c("Chris", "Charles", "Cameron", "Chloe", "Cora", "Caleb")) +
  labs(title = "Single observation - one time") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

p_mult_obs_flea <- ggplot() +
  theme_minimal() +
  theme_void() +
  coord_cartesian(xlim = c(-10, 8), ylim = c(-10, 10)) +
  geom_image(aes(x = c(-7, -7, -7), y = c(8, 0, -8), 
                 image = rep("images/flea_00.png", 3)), size = 0.3) +
  annotate("text", x = c(-7, -7, -7),
           y = c(8, 0, -8)-2.5, fontface = 2, size = 4.5, color = "gray50",
           label = c("Chris", "Charles", "Cameron")) +
  annotate("segment", x = rep(-3.5, 3), y = c(8, 0, -8), xend = rep(7.75, 3), yend = c(8, 0, -8),
           arrow = arrow(length = unit(0.03, "npc"), type = "closed")) +
  annotate("segment", x = rep(c(-1, 2, 5), times = 3), 
           y = c(7.8, 7.8, 7.8, -0.2, -0.2, -0.2, -8.2, -8.2, -8.2), 
           xend = rep(c(-1, 2, 5), times = 3), 
           yend = c(8.2, 8.2, 8.2, 0.2, 0.2, 0.2, -7.8, -7.8, -7.8)) +  
  annotate("text", x = rep(c(-1, 2, 5), times = 3), 
           y = c(7.8, 7.8, 7.8, -0.2, -0.2, -0.2, -8.2, -8.2, -8.2)-1,
           label = rep(c("t1", "t2", "t3"), times = 3), 
           fontface = 2, size = 4.5, color = "gray50") +
  labs(title = "Single observation - many times") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

geom_cage <- function(x, y) {
  list(
    geom_image(aes(x = c(x + 1, x + 1, x + 2.5, x + 4, x + 4), 
                   y = c(y + 1, y + 4, y + 2.5, y + 1, y + 4), 
                   image = rep("images/flea_00.png", 5)), size = 0.1),
    annotate("segment", x = x-0.25, y = y, xend = x + 5.25, yend = y),
    annotate("segment", x = x-0.25, y = y + 5, xend = x + 5.25, yend = y + 5),
    annotate("segment", x = x+seq(0, 5, by = 0.5), y = y, 
             xend = x +seq(0, 5, by = 0.5), yend = y + 5)
  )
}

p_block_obs_flea <- ggplot() +
  ## theme_minimal() +
  theme_void() +
  coord_cartesian(xlim = c(-8, 8), ylim = c(-8, 8.5)) +
  geom_cage(x = -7.5, y = -7.5) +
  geom_cage(x = -7.5, y = 2.5) +
  geom_cage(x = 2.5, y = -7.5) +
  geom_cage(x = 2.5, y = 2.5) +
  annotate("text", x = c(-7.5, 2.5, -7.5, 2.5) + 2.5,
           y = c(2.5, 2.5, -7.5, -7.5) + 6,
           label = c("Cage 1", "Cage 2", "Cage 3", "Cage 4"), 
           fontface = 2, size = 5, color = "gray50") +
  annotate("text", x = c(-7.5, 2.5, -7.5, 2.5) + 5.25,
           y = c(2.5, 2.5, -7.5, -7.5) - 0.75,
           label = c("n = 450", "n = 450", "n = 450", "n = 450"), 
           fontface = 4, size = 3.5, color = "gray50", hjust = "right") +
  labs(title = "Many observations combined") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

p_single_obs_flea + p_mult_obs_flea + p_block_obs_flea +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```



```{r}
jump_weight_tbl <- tibble(x = abs(rnorm(21, 3, 4)),
                          y = 10 + 1.2 * x + rnorm(21, 0, 2)) |>
  mutate_all(round, 1) |> 
  rename(weight = x, jumplength = y)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-01
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"

jump_weight_tbl |> 
  ggplot(aes(x = weight, y = jumplength, fill = col_pal(1, 0.5))) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21) +
  labs(x = "Cat flea weight [mg]", y = "Jump length in [cm]",
       fill = "Host") +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  scale_x_continuous(breaks = 0:10) +
  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
```

```{r}
library(mvtnorm)

sigma <- matrix(c(4,2,2,3), ncol=2)
x <- rmvnorm(n = 21, mean=c(1,2), sigma=sigma) |> 
  as_tibble() |> 
  rename(weight = V1, bodylength = V2)
colMeans(x)
var(x)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-02
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"

x |> 
  ggplot(aes(x = weight, y = bodylength, fill = col_pal(1, 0.5))) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21) +
  labs(x = "Cat flea weight [mg]", y = "Cat flea body length [mm]",
       fill = "Host") +
  scale_fill_viridis(discrete = TRUE, option = "plasma") #+
  # scale_x_continuous(breaks = 0:10) +
  # coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
```


### Non-linear


```{r}
jump_weight_non_linear_tbl <- tibble(x = abs(rnorm(32, 3, 4)),
                                     y = 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2 + rnorm(32, 0, 1)) |>
  rename(weight = x, jumplength = y)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-04
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"

jump_weight_non_linear_tbl |> 
  ggplot(aes(x = weight, y = jumplength, fill = col_pal(1, 0.5))) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21) +
  labs(x = "Cat flea weight [mg]", y = "Jump length in [cm]",
       fill = "Host") +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  scale_x_continuous(breaks = 0:10) +
  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
```


## Data availability

The data is available as txt-Files under <https://github.com/jkruppa/biodatascience>.

```{r}
#| echo: false

jump_weight_tbl |> 
  mutate_if(is.numeric, round, 2) |> 
  write_delim("data/jump_weight.txt")

jump_weight_non_linear_tbl |> 
  mutate_if(is.numeric, round, 2) |> 
  write_delim("data/jump_weight_non_linear.txt")
```


## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}