```{r echo = FALSE}
pacman::p_load(tidyverse, patchwork, knitr)
theme_book <- function() {
  theme_minimal() +
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(size = 16, face = "bold"),
          plot.subtitle = element_text(size = 12, face = "italic"),
          plot.caption = element_text(face = "italic"),
          axis.title = element_text(size = 12, face = "bold"),
          axis.text = element_text(size = 11),
          legend.title = element_text(face = "bold"))
}
```

# What is data? {#sec-what-is-data}

*Last modified on `r format(fs::file_info("chapter-02-what-is-data.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"The limits of my language mean the limits of my world." --- Ludwig Wittgenstein*

## General background

## Theoretical background


| Application | Symbol | Name | Description |
|:-------------|:--------:|:------------------|:------------------|
|  | $y$ | **outcome**, response, endpoint, dependent variable | *The right-hand side (abbr. RHS) of the model. Describing the values measured in an experiment or study.* |
|  | $x$ | **influencer**, influential variable, risk factor, fixed effect, independent variable | *The left-hand side (abbr. LHS) of the model. Describing the influential  variables in an experiment or study.* |
|  | $z$ | **random effect** | *A factor that provides a description of an grouping variable, which is not part of the controlled experimental setting.* |
| Explanation | $x$ | **explanator**, explanatory variable | *The influencer is used to describe or explain the outcome.* |
| Prediction | $x$ | **predictor**, predictive variable | *The influencer is used to predict the outcome.* |
| Main effect | $x$ | **focal explanator**, **focal predictor**, focal variable | *In a model with multiple influencers, the focal variable is the variable of primary interest.* |
| Continuous $x$ | $c$ | **covariate**, covariable | *The influencer is a numeric variable with continuous values.* |
| Categorical $x$ | $f_A$ | **factor A with levels A.1 to A.j**, factorial variable, categorical variable | *The influencer is discrete, functioning as a grouping variable, such as an experimental group or a treatment.*  |

: Table of terms used in the statistical modelling. The terms in bold are used here. Depending on the scientific background, the usage of these terms can vary widely.  {#tbl-names-regression tbl-colwidths="$$20, 5, 30, 45$$"}

A sentence why we use $y$ and not $x$ for mean and other stuff.

## R packages used

## Data

```{r}
jump_weight_tbl <- tibble(x = c(0.6, 1, 2.3, 3.5, 5.2, 7.1, 8.4, 9.2, 10),
                          y = 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2 + rnorm(9, 0, 0.5)) |>
  mutate_all(round, 1) |> 
  rename(weight_mg = x, jump_length_cm = y)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-02-jump-weight
#| tbl-cap: "foo."

jump_weight_tbl|> 
  kable(align = "c", "pipe")
```

@eq-jump-weight

$$
y = 0.15\cdot x^3 - 2.2\cdot x^2 + 8.8 \cdot x + 3.2
$$ {#eq-jump-weight}

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 10
#| fig-cap: "foo."
#| label: fig-02-science-to-model

p1 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Flea weight in [mg]", y = "Jump length in [cm]",
       title = "Science") +
  theme(axis.text = element_blank())

p2 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_point() +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "weight_mg", y = "jump_length_cm",
       title = "Data")

p3 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_function(fun = \(x) 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2,
                size = 1) +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Influencer (X)", y = "Outcome (Y)",
       title = "Model") +
  theme(axis.text = element_blank())

p1 + p2 + p3 +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 5.5
#| fig-cap: "foo."
#| label: fig-02-science-with-model

ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_point(size = 3, shape = 21, fill = "gray75") +
  geom_function(fun = \(x) 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2) +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Flea weight in [mg]", y = "Jump length in [cm]",
       title = "Science with data and model") 
```

## Alternatives

Further tutorials and R packages on XXX

## Glossary

term

:   what does it mean.

## The meaning of "Data Science" in this chapter.

-   itemize with max. 5-6 words

## Summary

## References {.unnumbered}
