```{r echo = FALSE, warning = FALSE}
source("init.R")
source("images/part_0/part_0_science.R")
```

# What is science? {#sec-what-is-science}

*Last modified on `r format(fs::file_info("chapter-01-what-is-science.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Reality is negotiable." --- Tim Ferriss*

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "foo"
#| label: fig-test-theory

c("#0D0887FF", "#6A00A8FF", "#B12A90FF", "#E16462FF", "#FCA636FF" ,"#F0F921FF")
c("#0D088780", "#6A00A880", "#B12A9080", "#E1646280", "#FCA63680", "#F0F92180")

ggplot() +
  #theme_void() +
  theme_minimal() +
  coord_cartesian(xlim = c(-10.5, 10.5), ylim = c(-10.5, 10.5)) +
  geom_hline(yintercept = c(-5, 0, 5), color = "gray") +
  geom_vline(xintercept = c(-5, 0, 5), color = "gray") +
  scale_x_continuous(breaks = seq(-10, 10, 1)) +
  scale_y_continuous(breaks = seq(-10, 10, 1)) +
  ## Population
  geom_shape(data = tibble(x = c(0.75, 10.25, 10.25, 0.75), 
                           y = c(0.75, 0.75, 9.75, 9.75)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "gray95") +
  geom_circle(aes(x0 = -5, y0 = 5, r = 4.25), fill = "white") +
  annotate("label", x = -5, y = 9.25, label = "Population", size = 7.5, fontface = 2) +
  annotate("text", x = 2, y = 9.95, label = "Randomisation", size = 5.5, fontface = 2) +
  annotate("text", x = 2, y = 9.5, label = "Causes structural equality", size = 5, fontface = 3) +
  annotate("text", x = 10.25, y = 5, label = "EXPERIMENT", size = 6, fontface = 2, angle = -90) +
  geom_curve(aes(x = -2, y = 8, xend = 6, yend = 8),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.3, color = "black", linewidth = 0.5) +
  ## Sample
  geom_circle(aes(x0 = 7.5, y0 = 7.5, r = 1.5), fill = "white") +
  annotate("label", x = 7.5, y = 9, label = "Sample", size = 7.5, fontface = 2) +
  annotate("text", x = 7.5, y = 7.5, label = "Data", size = 6, fontface = 2) +
  annotate("label", x = 8.75, y = 6.35, label = expression(D), size = 6) +
  geom_curve(aes(x = 6, y = 7, xend = 3, yend = 6),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = 0.4, color = "black", linewidth = 0.5) +
  geom_curve(aes(x = 9.1, y = 7.5, xend = 9.1, yend = 2.5),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.3, color = "black", linewidth = 0.5,
             linetype = "33") +
  ## Test
  geom_circle(aes(x0 = 7.5, y0 = 2.5, r = 1.5), fill = "white") +
  geom_curve(aes(x = 6, y = 2, xend = -1.9, yend = 2),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.3, color = "black", linewidth = 0.5) +
  annotate("text", x = 7.5, y = 2.25, label = "Statistical\ntest", size = 6, fontface = 2) +
  annotate("label", x = 8.75, y = 1.4, label = expression(T[D]), size = 6) +
  annotate("text", x = 2, y = 1.25, label = "Conclusion", size = 5.5, fontface = 2) +
  annotate("text", x = 2, y = 0.55, label = "Is the null hypothesis valid?", 
           size = 5, fontface = 3) +
  ## Desc stat
  geom_curve(aes(x = 3, y = 4, xend = 6, yend = 3),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = 0.4, color = "black", linewidth = 0.5) +
  geom_shape(data = tibble(x = c(1, 5, 5, 1), y = c(4.1, 4.1, 5.9, 5.9)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "white", color = "black") +
  annotate("text", x = 3, y = 5, label = "Descriptive\nstatistics", size = 6, fontface = 2) +
  ## Distribution
  geom_shape(data = tibble(x = c(-10.25, 10.25, 10.25, -10.25), 
                           y = c(-10.25, -10.25, -0.25, -0.25)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "gray95") +
  annotate("text", x = 10.25, y = -5, label = "THEORY", size = 6, fontface = 2, angle = -90) +
  geom_segment(aes(x = 7.5, y = -9, xend = 7.5, yend = -8)) +
  geom_line(data = tibble(x_raw = seq(-3.5, 3.5, 0.01), y_raw = dnorm(x_raw)), 
            aes(x = (x_raw*1.25 + 5), y = (y_raw*18 - 9)),
            linewidth = 1, color = "black") +
  geom_segment(aes(x = 5, y = -9, xend = 5, yend = -1.8), color = "black") +
  geom_segment(aes(x = 0.7, y = -9, xend = 9.3, yend = -9), color = "black") + 
  annotate("text", x = 7.5, y = -9.5, label = expression(T[D]), size = 6) +
  ## other samples
  geom_circle(aes(x0 = -9, y0 = -2, r = 1), fill = "white") +
  geom_circle(aes(x0 = -6.5, y0 = -3, r = 1), fill = "white") +
  geom_circle(aes(x0 = -3.5, y0 = -3, r = 1), fill = "white") +
  geom_circle(aes(x0 = -1, y0 = -2, r = 1), fill = "white") +
  geom_curve(aes(x = -3.5, y = 1, xend = -1.5, yend = -1),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = 0.2, color = "black", linewidth = 0.5) +
  geom_curve(aes(x = -4.5, y = 0.8, xend = -4, yend = -2),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = 0.2, color = "black", linewidth = 0.5) +
  geom_curve(aes(x = -5.5, y = 0.8, xend = -6, yend = -2),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.2, color = "black", linewidth = 0.5) +
  geom_curve(aes(x = -6.5, y = 1, xend = -8.5, yend = -1),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.2, color = "black", linewidth = 0.5)


```







```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-vulcan
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_vulcan_observed + p_vulcan_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-habital
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_habital_observed + p_habital_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-hollow
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"



p_hollow_observed + p_hollow_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

## General background

> *"Any sufficiently advanced technology is indistinguishable from magic." ---  @clarke1968clarke, Clarke's third law*

Full quote

> *"Reality is negotiable. Outside of science and law, all rules can be bent or broken, and it doesn’t require being unethical." --- Tim Ferriss*

Idea of hypotheses

Science is guessing and falsification

@chalmers2013thing [What is this thing called Science?](https://www.pmf.unizg.hr/_download/repository/1._Dodatna_literatura_%28filozofija_znanosti%29%5B3%5D.pdf)

@feynman1966science [What is science](https://www-oc.chemie.uni-regensburg.de/diaz/media/docs/feynman_what-is-science_68.pdf)

@campbell1952science [What is science?](https://www.whatisscience.info/files/what_is_science_rd1.pdf)

@clark2025models [Models Demystified: A Practical Guide from Linear Regression to Deep Learning](https://m-clark.github.io/book-of-models/)

@cox2017statistical [Statistical Thinking for the 21st Century](https://statsthinking21.github.io/statsthinking21-core-site/index.html)

@deutsch2011beginning [The beginning of infinity: Explanations that transform the world](https://www.academia.edu/download/106176299/rtx120400544p.pdf)

[David Deutsch > Quotes](https://www.goodreads.com/author/quotes/103190.David_Deutsch)

## Theoretical background

## R packages used

## Data

## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}