```{r echo = FALSE, warning = FALSE, message = FALSE}
source("init.R")
source("images/part_0/part_0_science.R")
```

# What is a model? {#sec-what-is-a-model}

*Last modified on `r format(fs::file_info("chapter-03-what-is-a-model.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

Imagine you are standing on a globe but still believe it is flat. You are standing on solid ground, with boiling rocks a few kilometers below you. A few kilometers above, you cannot breathe anymore. You are not moving, but the Earth is moving at speeds of up to 828,000 km/h, depending on the reference point. Your reality is a serious misinterpretation.

### Models in history

Johannes Kepler (1571–1630) was born into a world that believed the Earth was at the centre of the universe. The people of his time believed in a geocentric worldview. In 1543, Nikolaus Kopernikus (1473–1543) proposed a heliocentric model of the universe, with the Sun at the centre and the Earth orbiting it in a circular path. Nevertheless, the heliocentric worldview was still under debate when Kepler began his work. He studied the data observed by the astronomer Tycho Brahe. Brahe produced detailed observations of planetary movements. He made his observations without the use of a telescope. He was meticulous about the accuracy of his measurements of planetary orbits, and was one of the first people to recognise the importance of making multiple observations. In particular, Brahe observed the planet Mars. Kepler found that the positional data for Mars deviated by eight arc minutes from the Copernican circular orbit. This small deviation led Kepler to propose elliptical orbits. 

You might now be wondering what an arcminute is. It is a unit of measurement for degrees. Therefore, one degree has 60 arc minutes, and one arc minute has 60 arc seconds. If you look at the full moon, you will see that it covers 32 arc minutes. Kepler therefore believed that the eight-arc-minute deviation in the orbit of Mars was a systematic failure of the 2,000-year-old theory, rather than an error in the measurements taken by its mentor, Tycho Brahe.

Following the discovery of elliptical orbits, Sir Isaac Newton (1643–1727) formulated the Newtonian theory of gravity. This new theory enabled humanity to calculate gravity on Earth, the Moon's orbit around Earth, and the movement of the planets around the Sun. This is just one of many applications. Newton opened up a new field of classical physics. You might pause for a moment and ask yourself: if there is such a thing as classical physics, is there also modern physics? Yes and no. It depends on your perspective. Now, Newton became the benchmark for all new theories and models. Consequently, our observed data became increasingly detailed. We were able to determine the exact orbits of all planets. Everything seems to fit.

Problem solved. There was only one small deviation. The orbit of Mercury was slightly off. Specifically, the deviation was 43 arcseconds, or 1.2%, between the observed and calculated orbits using the Newtonian model. Albert Einstein (1879–1955) found this deviation intriguing. His General Theory of Relativity solved this mystery by merging time and space. Before Einstein's theory was formulated, astronomers suggested the existence of a new planet. They suggested that the deviation would go away if a planet called Vulcan existed and circled the sun in a very narrow orbit. However, this planet was never found.

Still Einstein believed that the solar system was in the middle of the galaxy. And there were no more galaxies than the milky way. 

Models can exists in parallel. Newtons laws of gravity can be used knowing that Newtons laws are wrong and the general relativity is less wrong. Why not right? This is science. 

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-vulcan
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_vulcan_observed + p_vulcan_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-habital
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_habital_observed + p_habital_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-hollow
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"



p_hollow_observed + p_hollow_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```


> *"Twistin', shake it, shake it, shake it, shake it, baby; Hey, we gonna Loop de Loop Shake it out, baby; Hey, we gonna Loop de La" --- [The Blues Brothers (1980) -- Shake a Tail Feather](https://www.youtube.com/watch?v=qdbrIrFxas0)*

## R packages used

## Data

## Alternatives

Further tutorials and R packages on XXX

## Glossary

term

:   what does it mean.

## The meaning of "Models of Reality" in this chapter.

-   itemize with max. 5-6 words

## Summary

## References {.unnumbered}
