```{r echo = FALSE, warning = FALSE, message=FALSE}
source("init.R")
```

# Visualisation of data 

*Last modified on `r format(fs::file_info("chapter-40-preface.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"What problem have you solved, ever, that was worth solving where you knew all the given information in advance? No problem worth solving is like that. In the real world, you have a surplus of information and you have to filter it, or you don't have sufficient information and you have to go find some." --- [Dan Meyer in  Math class needs a makeover](https://www.ted.com/talks/dan_meyer_math_class_needs_a_makeover?language=en&subtitle=en&trigger=5s)*

Here comes the preface text

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 7
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-reg-cross


geom_coord_cross <- function(x, y, yticks = "numeric", xticks = "numeric", zero = FALSE) {
  list(
    annotate("segment", x = x-0.2, xend = x+4, y = y, yend = y,
             arrow = arrow(length = unit(0.02, "npc"), type = "closed")),
    annotate("segment", x = x, xend = x, y = y-0.2, yend = y+4,
             arrow = arrow(length = unit(0.02, "npc"), type = "closed")),
    if(yticks == "numeric") {
      if(zero){
        list(
          annotate("segment", x = x-0.1, xend = x, y = y + seq(0, 3.5, 0.5), yend = y+ seq(0, 3.5, 0.5)),
          annotate("text", x = x -0.25, y = y + seq(0, 3.5, 0.5), label = seq(0, 70, 10),
                   hjust = "right"),
          annotate("text", x = x - 1, y = y +2, label = c("Count of hairleg"), fontface = 2, size = 5,
                   angle = 90)
        )
      } else {
        list(
          annotate("segment", x = x-0.1, xend = x, y = y + seq(0, 3.5, 0.5), yend = y+ seq(0, 3.5, 0.5)),
          annotate("text", x = x -0.25, y = y + seq(0.5, 3.5, 0.5), 
                   label = round(seq(15, 40, length.out = 7)), hjust = "right") ,
          annotate("text", x = x - 1, y = y +2, label = c("Jump length"), fontface = 2, size = 5, 
                   angle = 90)
        )        
      }
    },
    if(yticks == "3lvl") {
      list(
        annotate("segment", x = x-0.1, xend = x, y = y + (1:7)/2, yend = y + (1:7)/2),
        annotate("text", x = x -0.25, y = y + (1:7)/2, label = c("1", "2", "3", "4", "5", "6", "7"), 
                 hjust = "right") ,
        annotate("text", x = x - 1, y = y +2, label = c("Grade"), fontface = 2, size = 5, angle = 90)
      )
    },
    if(yticks == "2lvl") {
      list(
        annotate("segment", x = x-0.1, xend = x, y = y + c(0.5, 3), yend = y+  c(0.5, 3)),
        annotate("text", x = x -0.25, y = y + c(0.5, 3), label = c("0", "1"), hjust = "right") ,
        annotate("text", x = x - 1, y = y +2, label = c("Infected"), fontface = 2, size = 5, angle = 90)
      )
    },
    if(xticks == "2lvl") {
      list(
        annotate("segment", x = x+ c(0.5, 2.25), xend = x+ c(0.5, 2.25), y = y -0.1 , yend = y),
        annotate("text", x = x + c(0.5, 2.25), y = y -0.35, label = c("dog", "cat")),
        annotate("text", x = x + 2, y = y -1, label = c("Host"), fontface = 2, size = 5)

      )
    }
  )
}

ggplot() +
  ##theme_void() +
  theme_minimal() +
  coord_cartesian(xlim = c(2, 14), ylim = c(0, 14)) +
  annotate("segment", x = c(2, 8, 14), xend = c(2, 8, 14),
           y = c(0, 0, 0), yend = 14) +
  annotate("segment", x = c(0.25, 0.25, 0.25), xend = c(14, 14, 14),
           y = c(0, 7, 14), yend = c(0, 7, 14)) +
  scale_x_continuous(breaks = 0:14, expand = expansion(mult = 0)) +
  scale_y_continuous(breaks = 0:14, expand = expansion(mult = 0)) +
  ## outer text
  annotate("text", x = 11, y = 26.5, label = "Characteristic of the influencer (X)", 
           size = 8, fontface = 2) +
  annotate("text", x = c(5, 14), y = 25.25, 
           label = c("Covariate", "Factor"), size = 7, fontface = 2) +
  annotate("text", x = c(5, 11, 17), y = 24.5, 
           label = c("X is numeric", "X has 3 or more levels", "X has 2 levels"), 
           size = 6, fontface = 3) +
  ## coords
  geom_coord_cross(x = 3.5, y = 8.5, yticks = "numeric", xticks = "2lvl") +
  geom_coord_cross(x = 9.5, y = 8.5, yticks = "numeric", xticks = "2lvl", zero = TRUE) +
  geom_coord_cross(x = 3.5, y = 1.5, yticks = "3lvl", xticks = "2lvl") +
  geom_coord_cross(x = 9.5, y = 1.5, yticks = "2lvl", xticks = "2lvl") +
  annotate("text", x = c(11, 5, 11, 5), y = c(6.25, 6.25, 13.25, 13.25), 
           label = c("Binomial", "Ordinal", "Poisson", "Normal"), size = 7, fontface = 2) +
  ### normal
  geom_path(data = tibble(x_raw = seq(11-11/9, 
                                      11+11/9, 0.1), 
                          y_raw = dnorm(x_raw, mean = 11, sd = 0.4)), 
            aes(y = x_raw, x = y_raw*0.75+4.5), 
            linewidth = 1, color = "black", angle = 90) +
  geom_jitter(data = tibble(x = 4, y = rnorm(21, 11, 0.6)), aes(x, y), 
              shape = 21, width = 0.2, fill = "#0D088780") +
  geom_path(data = tibble(x_raw = seq(9.75-9.75/8.25, 
                                      9.75+9.75/8.25, 0.1), 
                          y_raw = dnorm(x_raw, mean = 9.75, sd = 0.4)), 
            aes(y = x_raw, x = y_raw*0.75+6.25), 
            linewidth = 1, color = "black", angle = 90) +
  geom_jitter(data = tibble(x = 5.75, y = rnorm(21, 9.75, 0.4)), aes(x, y), 
              shape = 21, width = 0.2, fill = "#A21D9A80") +
  ## Poisson
  geom_path(data = tibble(x_raw = seq(11-11/9, 
                                      11+11/9, 0.1), 
                          y_raw = dnorm(x_raw, mean = 11, sd = 0.4)), 
            aes(y = x_raw, x = y_raw*0.75+10.5), 
            linewidth = 1, color = "black", angle = 90) +
  geom_jitter(data = tibble(x = 10, y = (rpois(25, 3))/4+10.25), aes(x, y), 
              shape = 21, width = 0.25, height = 0, fill = "#0D088780") +
  geom_path(data = tibble(x_raw = seq(9.75-9.75/8.25, 
                                      9.75+9.75/8.25, 0.1), 
                          y_raw = dnorm(x_raw, mean = 9.75, sd = 0.4)), 
            aes(y = x_raw, x = y_raw*0.75+12.25), 
            linewidth = 1, color = "black", angle = 90) +
  geom_jitter(data = tibble(x = 11.75, y = (rpois(25, 3))/4+9), aes(x, y), 
              shape = 21, width = 0.25, height = 0, fill = "#A21D9A80") +
  ## ordinal
  geom_path(data = tibble(x_raw = seq(4-4/3.5, 
                                      4+4/3.5, 0.1), 
                          y_raw = dnorm(x_raw, mean = 4, sd = 0.4)), 
            aes(y = x_raw, x = y_raw*0.75+4.5), 
            linewidth = 1, color = "black", angle = 90) +
  geom_jitter(data = tibble(x = 4, y = ceiling(rnorm(21, 4, 0.4)/0.5)*0.5), aes(x, y), 
              shape = 21, width = 0.25, height = 0, fill = "#0D088780") +
  geom_path(data = tibble(x_raw = seq(2.75-2.75/2.5, 
                                      2.75+2.75/2.5, 0.1), 
                          y_raw = dnorm(x_raw, mean = 2.75, sd = 0.4)), 
            aes(y = x_raw, x = y_raw*0.75+6.25), 
            linewidth = 1, color = "black", angle = 90) +
  geom_jitter(data = tibble(x = 5.75, y = ceiling(rnorm(21, 2.75, 0.4)/0.5)*0.5), aes(x, y), 
              shape = 21, width = 0.25, height = 0, fill = "#A21D9A80") +
  ## binomial
  geom_path(data = tibble(x_raw = seq(3.875-3.875/3.5, 
                                      3.875+3.875/3.5, 0.1), 
                          y_raw = dnorm(x_raw, mean = 2+(2.5*0.75), sd = 0.4)), 
            aes(y = x_raw, x = y_raw*0.75+10.5), 
            linewidth = 1, color = "black", angle = 90) +
  geom_jitter(data = tibble(x = 10, y = (rbinom(21, 1, 0.75)*2.5)+2), aes(x, y), 
              shape = 21, width = 0.25, height = 0, fill = "#0D088780") +
  geom_path(data = tibble(x_raw = seq(2.625-2.625/2.5, 
                                      2.625+2.625/2.5, 0.1), 
                          y_raw = dnorm(x_raw, mean = 2+(2.5*0.25), sd = 0.4)), 
            aes(y = x_raw, x = y_raw*0.75+12.25), 
            linewidth = 1, color = "black", angle = 90) +
  geom_jitter(data = tibble(x = 11.75, y = (rbinom(21, 1, 0.25)*2.5)+2), aes(x, y), 
              shape = 21, width = 0.25, height = 0, fill = "#A21D9A80")  


jump_dog_tbl <- read_delim("data/jump_fleas.txt") |> 
  select(host, jumplength) |> 
  filter(host == "dog") 

mean_dog_jump <- mean(jump_dog_tbl$jumplength)
sd_dog_jump <- sd(jump_dog_tbl$jumplength)

 ggplot() +
  theme_book() +
  coord_cartesian(xlim = c(-0.25, 1.5), ylim = c(0, 60)) +
  geom_point(data = jump_dog_tbl, aes(x = 0, y = jumplength), size = 2) +
  geom_segment(aes(x = 0.75, xend = 0.75, 
               y = mean_dog_jump, yend = mean_dog_jump+ sd_dog_jump), 
               color = "#E16462FF", size = 1) +
  annotate("text", x = 0.8, y = mean_dog_jump+ sd_dog_jump/2-0.5, hjust = "left", 
           label = expression(bold(s)), size = 5, color = "#E16462FF") +
  geom_path(data = tibble(x_raw = seq(mean_dog_jump-mean_dog_jump/1.5, 
                                      mean_dog_jump+mean_dog_jump/1.5, 0.1), 
                          y_raw = dnorm(x_raw, mean = mean_dog_jump, sd = sd_dog_jump)), 
            aes(y = x_raw, x = y_raw*20), 
            linewidth = 1, color = "black", angle = 90) +
  geom_segment(aes(x = -0.03, xend = 1.35, 
               y = mean_dog_jump, yend = mean_dog_jump), color = "#B12A90FF", size = 1) +
  annotate("text", x = 1.4, y = 33, hjust = "left", label = expression(bold(bar(y))),
           size = 5, color = "#B12A90FF") + 
  geom_text_repel(data = jump_dog_tbl, aes(x = 0, y = jumplength, label = 1:7)) +
  scale_x_continuous(breaks = 0, labels = "dog") +
  scale_y_continuous(breaks = c(0, 20, mean_dog_jump, 40, 60),
                     labels = c("0", "20", round(mean_dog_jump,1), "40", "60")) +
  labs(x = "", y = "")

```

