```{r echo = FALSE, warning = FALSE, message = FALSE}
source("init.R")
source("images/part_0/part_0_data.R")
```

# What is data? {#sec-what-is-data}

*Last modified on `r format(fs::file_info("chapter-02-what-is-data.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"The limits of my language mean the limits of my world." --- Ludwig Wittgenstein*

Emma sat on the wooden stage. She need time to recover from what she had seen behind the curtains. What she had seen beneath the wood had now become reality. Dust danced in the light. Single light beams hit some dust particles from time to time and exploded into golden rain. Sparkles dropped on the wood.

"How is gold made?" Emma asked.

"This was a long break. I'm a bit surprised by your question. I thought we should discuss your findings from behind the stage. Not gold." Jeff answers from his seat on the left. 

"Gold cannot be formed through fusion in a star. This is only possible up to iron. You need a supernova or a collision of two neutron stars." Emma spoke absently into the darkness.

"Interesting topic. Was this observed? The collision of two neutron stars and the resulting gold spewing out," Jeff laughed.

"No, it was a computer simulation. Artificial data, if you will." Emma looked around. She decided that it was time to begin the play.

"Then it's not real, and it's not an explanation for me at all." Jeff stated.

"But the simulation was based on real observations." Emma stood up and started building the scene for the play.

## What is observed?

Can we observe anything? What does it mean to observe? We see things, but our vision is already influenced by our understanding of what they are. Additionally, by naming things and processes, we are already following a theory. "The sun is rising" is a popular example of a saying connected with the geocentric theory, which states that the Earth is at the center of the solar system. Therefore, our language is already influenced by our observations. We cannot speak about something we observe without giving the observation some meaning. It is possible, but it demands a high level of concentration.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 0.75
#| fig-width: 7.5
#| fig-cap: "An observation is a measurement or sensory input within a theoretical and linguistic framework. An observation cannot be made without a context. There is no such thing as an objective observation."
#| label: fig-observation

p_observation_data
```

This book requires a highly specific type of observation. We need data. Very rarely is any observation we make based on first-hand data. For example, if you look outside your window and observe a bird, you will learn some information about it. But is this data? No, it is just information. Data is information, but structured information. Data is a type of observation that is stored in a specific way. We translate observations and information about them into a set of numbers. Not every observation can be represented as a number in a data sheet or table. As their names suggest, data sheets and tables are two-dimensional. A sheet is a type of paper on which organised observations can be written down. The same is true for a table. It is a structured type of observation.

## What is data?

We will now think about data. This will happen after we have discussed the science in the chapter. It will also happen if you have skipped the chapter entirely. What is data? If you had asked me when I was a teenager, I would have referred to Star Trek: The Next Generation and Commander Data. He was an android without any emotions. Pure objectivity. He was an artificial being. However, this is not the type of data we want to discuss. We want to discuss data that has been observed and measured. Furthermore, our data is numeric or has a numeric counterpart. We use descriptive words, but rarely. Later on, we want to run algorithms and calculations with our data. Therefore, the data must be open to calculations.

In this book, we will follow an inductive approach. First, we make observations, then we try to find patterns in them. Therefore, our data is the fuel of discovery. Conversely, we can also use data to test a hypothesis. Both are valid and are used. Since this book is about statistics and data science, we cannot generate new theories. We are limited by the tools we have and the data available to us. This book may be read by those with a background in natural science. Then, you can use the statistical tools differently.

Can we use data in any format? No, you cannot. There is a general format for how data should be stored. In our case, we have observations in the rows and variables of measurement in the columns. There can only be one observation per row. This format is called the long format. Another possibility is the wide format, where the observations are scattered around and do not share the same rows. The wide format may be useful for time series, but I do not recommend it.

Data are observations represented by numbers and letters.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.25
#| fig-width: 7
#| fig-cap: "foo. **(A)** foo. **(B)**"
#| label: fig-02-data

ggplot() +
  theme_void() +
  ##theme_minimal() +
  coord_cartesian(xlim = c(-0.45, 6), ylim = c(0, 5.8)) +
  scale_x_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  geom_shape(data = tibble(x = c(1.1, 5.9, 5.9, 1.1), 
                           y = c(0.1, 0.1, 3.9, 3.9)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "gray95") +
  annotate("segment", x = 0, xend = 10, y = 1:4, yend = 1:4) +
  annotate("segment", x = 1:5, xend = 1:5, y = 0, yend = 5) +
  annotate("text", x = seq(1.5, 4.5, 1), y = 4.5,
           label = c(1, 2, 3, "..."), fontface = 2, size = 5) +
  annotate("text", x = 5.5, y = 4.5, label = "p", fontface = 2, size = 5) +
  annotate("text", y = seq(3.5, 1.5, -1), x = 0.5,
           label = c(1, 2, "..."), fontface = 2, size = 5) +
  annotate("text", x = 0.5, y = 0.5, label = "n", fontface = 2, size = 5) +
  annotate("segment", x = 0, xend = 1, y = 5, yend = 4) +
  annotate("text", x = 0.25, y = 4.3, label = "[i, ]", fontface = 2, size = 5) +
  annotate("text", x = 0.8, y = 4.75, label = "[ , j]", fontface = 2, size = 5) +
  annotate("text", x = 3.5, y = 5.5, label = "Columns", size = 7, fontface = 2) +
  annotate("text", x = -0.25, y = 2.5, label = "Rows", size = 7, fontface = 2, angle = 90) +
  annotate("text", x = c(seq(1.5, 3.5, 1), 5.5), y = 1.5, label = "...") +
  annotate("text", x = 4.5, y = c(0.5, 2.5, 3.5), label = "...") +
  annotate("text", x = c(1.5, 2.5, 3.5, 5.5, 1.5, 2.5, 3.5, 5.5, 1.5, 2.5, 3.5, 5.5, 4.5),
           y = c(3.5, 3.5, 3.5, 3.5, 2.5, 2.5, 2.5, 2.5, 0.5, 0.5, 0.5, 0.5, 1.5),
           label = c("[1, 1]", "[1, 2]", "[1, 3]", "[1, p]",
                     "[2, 1]", "[2, 2]", "[2, 3]", "[2, p]",
                     "[n, 1]", "[n, 2]", "[n, 3]", "[n, p]", "[i, j]"), size = 4.5)
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.25
#| fig-width: 7
#| fig-cap: "foo. **(A)** foo. **(B)**"
#| label: fig-02-data-complex

ggplot() +
  theme_void() +
  #theme_minimal() +
  coord_cartesian(xlim = c(-0.45, 6), ylim = c(0, 5.8)) +
  scale_x_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  geom_shape(data = tibble(x = c(0.1, 0.9, 0.9, 0.1), 
                           y = c(0.1, 0.1, 4.6, 4.6)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "#6A00A880", alpha = 0.2) +
  geom_shape(data = tibble(x = c(1.1, 2.9, 2.9, 1.1), 
                           y = c(0.1, 0.1, 4.6, 4.6)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "#B12A9080", alpha = 0.2) +
  geom_shape(data = tibble(x = c(3.1, 5.9, 5.9, 3.1), 
                           y = c(0.1, 0.1, 4.6, 4.6)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "#FCA63680", alpha = 0.2) +
  annotate("segment", x = 0, xend = 10, y = 1:4, yend = 1:4) +
  annotate("segment", x = 1:5, xend = 1:5, y = 0, yend = c(5, 4.6, 5, 4.6, 4.6)) +
  annotate("text", x = seq(0.5, 5.5, 1), y = 4.3,
           label = c(expression(bold(ID)), expression(bold(X[1])), expression(bold(X[2])), 
                     expression(bold(Y[1])), expression(bold(Y[2])), expression(bold(Y[3]))), 
           fontface = 2, size = 5) +
  annotate("label", x = c(0.5, 2, 4.5), y = 4.9,
           label = c("Individual", "Experimental", "Measurement"),
           fill = c("#6A00A880", "#B12A9080", "#FCA63680"), size = 4, fontface = 2) +
  annotate("text", y = seq(3.5, 1.5, -1), x = 0.5,
           label = c(1, 2, "..."), fontface = 2, size = 5) +
  annotate("text", x = 0.5, y = 0.5, label = "n", fontface = 2, size = 5) +
  annotate("text", x = 3.5, y = 5.5, label = "Columns", size = 7, fontface = 2) +
  annotate("text", x = -0.25, y = 2.5, label = "Rows", size = 7, fontface = 2, angle = 90) +
  annotate("text", x = rep(c(1.5, 2.5, 3.5, 4.5, 5.5), each = 4),
           y = rep(c(0.5, 1.5, 2.5, 3.5), times = 5), 
           label = c(10.1, 12.4, 15.6, 22.4, "cat", "cat", "cat", "dog", 
                     2, 2, 4, 1, 22.1, 19.7, 24.3, 33.2, 1, 1, 0, 1),
           size = 4.5)
```

What is with time series?


@hassenstein2025ten

@wilkinson2016fair

## Simulation and artificial data
Artificial data is model-laden

## Technical vs. biological replication

## Experimental vs. observed data

## Clinical studies

## Long format vs. wide format

## The elephant in the room is Excel

## Tidy data

## Theoretical background

## R packages used

Do we want here to simulate data?

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3.25
#| fig-width: 7
#| fig-cap: "foo. **(A)** foo. **(B)**"
#| label: fig-02-cabin-tibble

p_tibble_empty <- ggplot() +
  theme_void() +
 # theme_minimal() +
  coord_cartesian(xlim = c(0, 10), ylim = c(0.5, 10.5)) +
  scale_x_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  geom_cabin(x = 0.5, y = 0.5) +
  labs(title = "A tibble()") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic")) 

p_tibble_filled <- ggplot() +
  theme_void() +
 # theme_minimal() +
  coord_cartesian(xlim = c(0, 10), ylim = c(0.5, 10.5)) +
  scale_x_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  geom_cabin(x = 0.5, y = 0.5) +
  geom_hook(x = 1.7, y = 8.25, width = 1.3, angle = 0,
            hook_text = str_c(c("cat", "cat", "cat", "...", "dog", "dog", "dog"), 
                              collapse = "\n"), label = FALSE) +
  geom_hook(x = 3.7, y = 8.25, width = 1.3, angle = 0,
            hook_text = str_c(c("33.2", "31.7", "36.8", "...", "22.1", "19.7", "24.3"), 
                              collapse = "\n"), label = FALSE) +
  geom_hook(x = 5.7, y = 8.25, width = 1.3, angle = 0,
            hook_text = str_c(c("10.1", "12.4", "15.6", "...", "22.4", "18.1", "25.6"), 
                              collapse = "\n"), label = FALSE) +
  geom_hook(x = 7.7, y = 8.25, width = 1.3, angle = 0,
            hook_text = str_c(c("yes", "yes", "no", "...", "no", "yes", "no"), 
                              collapse = "\n"), label = FALSE) +
  labs(title = "A tibble()") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 12, face = "italic"))

p_tibble_empty + p_tibble_filled +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 3
#| fig-width: 4
#| fig-cap: "foo."
#| label: fig-02-cabbinet

ggplot() +
  theme_void() +
 # theme_minimal() +
  coord_cartesian(xlim = c(0, 10), ylim = c(0.5, 10)) +
  scale_x_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(breaks = seq(-10,10,1), expand = expansion(mult = c(0, 0))) +
  geom_cabin(x = 0.5, y = 0.5) +
  geom_hook(x = 1.5, y = 8.25, width = 0.8, angle = 0,
            hook_text = str_c(c(2.1, 3.2, 2.1, 3.2, 2.1, 3.2, 2.1, 3.2), collapse = "\n"), 
            label_text = "hust") +
  geom_hook(x = 2.75, y = 8.25, width = 0.8, angle = 0,
            hook_text = str_c(c(2.1, 3.2, 2.1, 3.2, 2.1, 3.2, 2.1, 3.2), collapse = "\n"), 
            label_text = "hust") +
  geom_hook(x = 4, y = 8.25, width = 0.8, angle = 0,
            hook_text = str_c(c(2.1, 3.2, 2.1, 3.2, 2.1, 3.2, 2.1, 3.2), collapse = "\n"), 
            label_text = "hust") +
  geom_hook(x = 5.25, y = 8.25, width = 0.8, angle = 0,
            hook_text = str_c(c(2.1, 3.2, 2.1, 3.2, 2.1, 3.2, 2.1, 3.2), collapse = "\n"), 
            label_text = "hust") +
  geom_hook(x = 5.25, y = 8.25, width = 0.8, angle = 0,
            hook_text = str_c(c(2.1, 3.2, 2.1, 3.2, 2.1, 3.2, 2.1, 3.2), collapse = "\n"), 
            label_text = "hust") +
  geom_hook(x = 6.5, y = 8.25, width = 0.8, angle = 0,
            hook_text = str_c(c(2.1, 3.2, 2.1, 3.2, 2.1, 3.2, 2.1, 3.2), collapse = "\n"), 
            label_text = "hust") +
  geom_hook(x = 7.75, y = 8.25, width = 0.8, angle = 0,
            hook_text = str_c(c(2.1, 3.2, 2.1, 3.2, 2.1, 3.2, 2.1, 3.2), collapse = "\n"), 
            label_text = "hust")
```


## Data

Is your data the fuel of discovery or the test of discovery?

```{r}
jump_weight_tbl <- tibble(x = c(0.6, 1, 2.3, 3.5, 5.2, 7.1, 8.4, 9.2, 10),
                          y = 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2 + rnorm(9, 0, 0.5)) |>
  mutate_all(round, 1) |> 
  rename(weight_mg = x, jump_length_cm = y)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-02-jump-weight
#| tbl-cap: "foo."

jump_weight_tbl|> 
  kable(align = "c", "pipe")
```

@eq-jump-weight

$$
y = 0.15\cdot x^3 - 2.2\cdot x^2 + 8.8 \cdot x + 3.2
$$ {#eq-jump-weight}

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 10
#| fig-cap: "foo."
#| label: fig-02-science-to-model

p1 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Flea weight in [mg]", y = "Jump length in [cm]",
       title = "Science") +
  theme(axis.text = element_blank())

p2 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_point() +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "weight_mg", y = "jump_length_cm",
       title = "Data")

p3 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_function(fun = \(x) 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2,
                size = 1) +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Influencer (X)", y = "Outcome (Y)",
       title = "Model") +
  theme(axis.text = element_blank())

p1 + p2 + p3 +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 5.5
#| fig-cap: "foo."
#| label: fig-02-science-with-model

ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_point(size = 3, shape = 21, fill = col_pal(1, 0.5)) +
  geom_function(fun = \(x) 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2) +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Flea weight in [mg]", y = "Jump length in [cm]",
       title = "Science with data and model") 
```


## Outro

What do we want to do in this book? We want to perform inductive reasoning based on data. Statistical modeling is inductive reasoning, so it is also based on data. Therefore, we will need data to perform statistical analyses in any possible combination. In other words, we live in a data-driven world.


## Alternatives

Further tutorials and R packages on XXX

## Glossary

term

:   what does it mean.

## The meaning of "Models of Reality" in this chapter.

-   itemize with max. 5-6 words

## Summary

## References {.unnumbered}
