```{r echo = FALSE, warning = FALSE}
source("init.R")
source("images/part_3/part_3_theory_testing.R")

```

# Statistical testing {#sec-temp}

*Last modified on `r format(fs::file_info("chapter-31-statistical-testing.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

## General background



```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-test-theory-pval-fisher

p_val_tbl <- expand_grid(x = 1:50,
                         y = 1:20) |> 
  mutate(fisher = sort(round(pt(rt(1000, 5), 5), 3)),
         newman = as.numeric(fisher >= 0.05)) |> 
  pivot_longer(cols = fisher:newman)


ggplot() +
  theme_void() +
  ##theme_minimal() +
  coord_cartesian(xlim = c(-2, 50), ylim = c(-2, 47)) +
  geom_shape(data = tibble(x = c(-3, 52, 52, -3), 
                           y = c(-4, -4, 24.25, 24.25)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "gray95") +
  geom_shape(data = tibble(x = c(-3, 52, 52, -3), 
                           y = c(24.75, 24.75, 49.5, 49.5)),
             aes(x, y), radius = unit(0.5, 'cm'), fill = "gray95") +
  annotate("text", x = -3, y = 37, label = "FISHER", size = 5.5, fontface = 2, 
           angle = 90) +
  annotate("text", x = -3, y = 10.5, label = "NEWMAN/PEARSON", size = 5.5, fontface = 2, 
           angle = 90) +
  geom_point(data = filter(p_val_tbl, name == "newman"),
             aes(x, y, fill = value),
             shape = 21, size = 3, alpha = 0.75) +
  geom_point(data = filter(p_val_tbl, name == "fisher"),
             aes(x, y+25, fill = value),
             shape = 21, size = 3, alpha = 0.75) +
  geom_curve(aes(x = 49, y = 47, xend = 2, yend = 47),
             arrow = arrow(length = unit(0.02, "npc"), type = "closed"),
             curvature = 0, color = "black", linewidth = 1) +
  annotate("label", hjust = "center", x = 25, y = 47.15,
           label = "Less likely to observe your data under the Null",
           fontface = 2, size = 4) +
  annotate("label", hjust = "left", x = 10, y = 22.5,
           label = expression(alpha~"="~5*"%")) +
  geom_curve(aes(x = 9.8, y = 22.5, xend = 3, yend = 10.5),
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = 0.3, color = "black", linewidth = 0.75) +
  annotate("label", hjust = "center", x = 2, y = 22.5,
           label = expression(p <= alpha), fill = "#0D088780") +
  annotate("label", hjust = "center", x = 27, y = 22.5,
           label = expression(p > alpha), fill = "#F0F92180") +
  annotate("label", hjust = "center", x = 27, y = -1.5,
           label = "Accept Null", fontface = 2, fill = "#F0F92180", size = 3.5) +
  annotate("label", hjust = "center", x = 2, y = -1.5,
           label = "Reject Null", fontface = 2, fill = "#0D088780", size = 3.5) +
  scale_fill_viridis(option = "plasma") +
  theme(legend.position = "none") 
```



```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-test-theory-dist
p_population_sample
```


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 10
#| fig-cap: "foo"
#| label: fig-test-theory-random
p_theory_random
```



```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4.75
#| fig-width: 10
#| fig-cap: "foo"
#| label: fig-test-theory-upper
p_theory_random_upper
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "foo"
#| label: fig-test-theory-full
p_theory_random_full
```

## Theoretical background

### Fisher's approach: The ‘significance test’

Fisher saw statistics as a tool for inductive reasoning (learning from data for science).

-   Only ONE hypothesis: There is only the null hypothesis ($H_0$ e.g., ‘no effect’). An alternative does not formally exist.
-   The measure (p-value): The p-value is a continuous measure of the strength of evidence against the null hypothesis $H_0$.
    -   $p = 0.01$ Strong evidence against the null hypothesis.
    -   $p = 0.20$ No evidence against the null hypothesis. 
-   The result: One rejects the null hypothesis $H_0$ or one does not make a judgement. One never ‘accepts’ the null hypothesis (one simply has not found enough evidence to reject it).
-   Objective: Gain knowledge through individual experiments.

### Neyman-Pearson's approach: The ‘hypothesis test’

Neyman and Pearson sharply criticised Fisher.
They said, ‘You can't reject anything if you don't know what to accept instead.’ They saw statistics as a decision-making process (behaviourism).

-   TWO hypotheses: There is the null hypothesis ($H_0$) AND a specific alternative hypothesis ($H_A$).
-   Type 1 and 2 errors: Before the experiment begins, the following is determined:
    -   $\alpha$ (alpha): How often am I allowed to incorrectly find an effect? (e.g. 5%)
    -   $\beta$ (Beta): How often am I allowed to mistakenly overlook a real effect? (Power/test strength).
-   The result: A tough decision. ‘Accept $H_0$’ or ‘Reject $H_0$’ (or Accept $H_A$).
-   Goal: Minimisation of losses over many repeated experiments (as in industrial production).

Neymans philosophy: We are not looking for the ‘truth’ in individual cases, but rather we behave in such a way that we are wrong as rarely as possible in 1000 decisions.

### Today's ‘hybrid chaos’

Modern textbooks and software (such as SPSS or R) often use a hybrid that historically makes no sense:

-   We define $\alpha = 5\%$ (Neyman-Pearson).
-   We calculate an exact p-value (Fisher).
-   We report the p-value as evidence (Fisher), but use it for a hard yes/no decision (Neyman-Pearson).
-   We talk about ‘power’ (Neyman-Pearson), but often only test against a non-specific alternative.

This mishmash often leads to misunderstandings, such as that a $p = 0.001$ indicates a ‘stronger effect’ than $p = 0.049$  (Fisher thinking), even though in Neyman-Pearson logic at $\alpha = 5\%$ , one would have to make exactly the same decision in both cases (‘Reject $H_0$’).

## R packages used

## Data

## Alternatives

Further tutorials and R packages on XXX

## Glossary

term

:   what does it mean.

## The meaning of "Models of Reality" in this chapter.

-   itemize with max. 5-6 words

## Summary

## References {.unnumbered}
