```{r echo = FALSE, warning = FALSE, message = FALSE}
source("init.R")
source("images/part_0/part_0_science.R")
```

# What is a model? {#sec-what-is-a-model}

*Last modified on `r format(fs::file_info("chapter-03-what-is-a-model.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

Imagine you are standing on a globe but still believe it is flat. You are standing on solid ground, with boiling rocks a few kilometers below you. A few kilometers above, you cannot breathe anymore. You are not moving, but the Earth is moving at speeds of up to 828,000 km/h, depending on the reference point. Your reality is a serious misinterpretation.


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-vulcan
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_vulcan_observed + p_vulcan_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-habital
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_habital_observed + p_habital_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-hollow
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"



p_hollow_observed + p_hollow_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```


> *"Twistin', shake it, shake it, shake it, shake it, baby; Hey, we gonna Loop de Loop Shake it out, baby; Hey, we gonna Loop de La" --- [The Blues Brothers (1980) -- Shake a Tail Feather](https://www.youtube.com/watch?v=qdbrIrFxas0)*

## R packages used

## Data

## Alternatives

Further tutorials and R packages on XXX

## Glossary

term

:   what does it mean.

## The meaning of "Models of Reality" in this chapter.

-   itemize with max. 5-6 words

## Summary

## References {.unnumbered}
