```{r echo = FALSE, warning = FALSE}
source("init.R")
source("images/part_0/part_0_science.R")
```

# What is science? {#sec-what-is-science}

*Last modified on `r format(fs::file_info("chapter-01-what-is-science.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Reality is negotiable." --- Tim Ferriss*


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 7
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-test-theory
set.seed(20240407)

truth_tbl <- tibble(grp = as_factor("truth"),
                    x = rnorm(200, 20, 7),
                    y = rnorm(200, 70, 7),
                    color = as_factor(sample(1:6, 200, replace = TRUE)),
                    shape = as_factor(sample(1:4, 200, replace = TRUE)))

data_theory_tbl <- tibble(grp = as_factor("data"),
                          x = rnorm(12, 90, 2),
                          y = rnorm(12, 80, 2),
                          color = as_factor(sample(1:3, 12, replace = TRUE)),
                          shape = as_factor(sample(1:2, 12, replace = TRUE)))

t1_tbl <- tibble(grp = as_factor("t1"),
                 x = rnorm(12, 70, 3),
                 y = rnorm(12, 45, 3),
                 color = as_factor(sample(2:4, 12, replace = TRUE)),
                 shape = as_factor(sample(2:3, 12, replace = TRUE)))

t2_tbl <- tibble(grp = as_factor("t1"),
                 x = rnorm(12, 60, 2),
                 y = rnorm(12, 25, 2),
                 color = as_factor(sample(4:6, 12, replace = TRUE)),
                 shape = as_factor(sample(2:3, 12, replace = TRUE)))

t3_tbl <- tibble(grp = as_factor("t1"),
                 x = rnorm(12, 35, 2),
                 y = rnorm(12, 10, 2),
                 color = as_factor(sample(1:4, 12, replace = TRUE)),
                 shape = as_factor(sample(3:4, 12, replace = TRUE)))

t4_tbl <- tibble(grp = as_factor("t1"),
                 x = rnorm(12, 5, 3),
                 y = rnorm(12, 10, 3),
                 color = as_factor(sample(3:6, 12, replace = TRUE)),
                 shape = as_factor(sample(1:2, 12, replace = TRUE)))

dist_tbl <- tibble(x_raw = seq(-4, 4, 0.01),
                   y_raw = dnorm(x_raw))

bind_rows(truth_tbl,
          data_theory_tbl,
          t1_tbl,
          t2_tbl,
          t3_tbl,
          t4_tbl) |> 
  ggplot(aes(x, y)) +
  theme_void() +
  #theme_minimal() +
  geom_jitter(aes(fill = color, shape = shape), size = 4, width = 5, height = 5) +
  annotate("text", x = 60, y = 105, label = "Randomisierung", size = 5, fontface = 2) +
  annotate("text", x = 60, y = 101, label = "(Strukturgleichheit)", size = 5, fontface = 3) +
  geom_curve(x = 36, y = 90, xend = 85, yend = 85,
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.5, color = "black", linewidth = 0.5) +
  geom_curve(x = 91, y = 38, xend = 43, yend = 72,
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = 0.6, color = "black", linewidth = 0.5) +
  geom_curve(x = 99, y = 83, xend = 110, yend = 68,
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.2, color = "black", linewidth = 0.5) +
  geom_curve(x = 110, y = 60, xend = 105, yend = 11,
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.3, color = "black", linewidth = 0.5) +
  annotate("text", x = 20, y = 102, label = "Grundgesamtheit", size = 5, fontface = 2) +
  geom_ellipse(aes(x0 = 20, y0 = 70, a = 22, b = 29, angle = 0), color = "black") +
  annotate("text", x = 93, y = 98, label = "Stichprobe", size = 5, fontface = 2) +
  annotate("text", x = 93, y = 94, label = "Daten (D)", size = 5, fontface = 3) +
  geom_ellipse(aes(x0 = 91, y0 = 80, a = 12, b = 7, angle = 20), color = "black") +
  annotate("text", x = 110, y = 66, label = "Teststatistik", size = 5, fontface = 2) +
  annotate("label", x = 110, y = 60, label = expression(T[D]), size = 7) +
  annotate("text", x = 60, y = 77, label = "Rückschluß", size = 5, fontface = 2) +
  annotate("text", x = 60, y = 70, label = expression(Gilt~die~H[0]*"?"), 
           size = 5, fontface = 3) +
  geom_ellipse(aes(x0 = 7, y0 = 10, a = 10, b = 10, angle = 0), color = "gray") +
  geom_ellipse(aes(x0 = 36, y0 = 10, a = 10, b = 10, angle = 0), color = "gray") +
  geom_ellipse(aes(x0 = 60, y0 = 24, a = 10, b = 10, angle = 0), color = "gray") +
  geom_ellipse(aes(x0 = 70, y0 = 44, a = 10, b = 10, angle = 0), color = "gray") +
  geom_curve(x = 34, y = 42, xend = 59, yend = 44,
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.2, color = "gray", linewidth = 0.5) +
  geom_curve(x = 35, y = 41, xend = 51, yend = 30,
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.2, color = "gray", linewidth = 0.5) +
  geom_curve(x = 30, y = 40, xend = 35, yend = 21,
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = -0.2, color = "gray", linewidth = 0.5) +
  geom_curve(x = 20, y = 40, xend = 8, yend = 21,
             arrow = arrow(length = unit(0.01, "npc"), type = "closed"),
             curvature = 0.2, color = "gray", linewidth = 0.5) +
  annotate("label", x = 25, y = 42, label = expression(H[0]~wahr), 
           size = 7) +
  geom_line(data = dist_tbl, aes(x = (x_raw + 13)*7, y = y_raw*80+5),
            linewidth = 1, color = "gray") +
  geom_segment(x = 65, y = 5, xend = 115, yend = 5, color = "gray") +
  geom_segment(x = 91, y = 5, xend = 91, yend = 37, color = "gray") +
  geom_segment(x = 104, y = 5, xend = 104, yend = 10.5) +
  annotate("text", x = 91, y = 3, label = "0") +
  annotate("text", x = 104, y = 3, label = expression(T[D])) +
  annotate("text", x = 80, y = -2, label = "Testverteilung", size = 5, 
           fontface = 2) +
  annotate("text", x = 101, y = -2, label = expression("("*T[1]*","*"..."*","*T[p]*")"), size = 5, 
           fontface = 2) +
  annotate("text", x = 21.5, y = 12, label = "...", size = 12, fontface = 2,
           color = "gray") +
  annotate("label", x = 81, y = 42, label = expression(T[1]), size = 7) +
  annotate("label", x = 70, y = 18, label = expression(T[2]), size = 7) +
  annotate("label", x = 46, y = 3, label = expression(T[3]), size = 7) +
  annotate("label", x = 16.5, y = 3, label = expression(T[p]), size = 7) +
  scale_x_continuous(limits = c(-3, 115)) +
  scale_y_continuous(limits = c(-4, 105)) +
  scale_shape_manual(values = 21:24) +
  theme(legend.position = "none") +
  scale_fill_okabeito() 
```



```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-vulcan
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_vulcan_observed + p_vulcan_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-habital
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_habital_observed + p_habital_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-hollow
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"



p_hollow_observed + p_hollow_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

## General background

> *"Any sufficiently advanced technology is indistinguishable from magic." ---  @clarke1968clarke, Clarke's third law*

Full quote

> *"Reality is negotiable. Outside of science and law, all rules can be bent or broken, and it doesn’t require being unethical." --- Tim Ferriss*

Idea of hypotheses

Science is guessing and falsification

@chalmers2013thing [What is this thing called Science?](https://www.pmf.unizg.hr/_download/repository/1._Dodatna_literatura_%28filozofija_znanosti%29%5B3%5D.pdf)

@feynman1966science [What is science](https://www-oc.chemie.uni-regensburg.de/diaz/media/docs/feynman_what-is-science_68.pdf)

@campbell1952science [What is science?](https://www.whatisscience.info/files/what_is_science_rd1.pdf)

@clark2025models [Models Demystified: A Practical Guide from Linear Regression to Deep Learning](https://m-clark.github.io/book-of-models/)

@cox2017statistical [Statistical Thinking for the 21st Century](https://statsthinking21.github.io/statsthinking21-core-site/index.html)

@deutsch2011beginning [The beginning of infinity: Explanations that transform the world](https://www.academia.edu/download/106176299/rtx120400544p.pdf)

[David Deutsch > Quotes](https://www.goodreads.com/author/quotes/103190.David_Deutsch)

## Theoretical background

## R packages used

## Data

## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}