```{r echo = FALSE}
pacman::p_load(tidyverse)
theme_book <- function() {
  theme_minimal() +
    theme(panel.grid.minor = element_blank(),
          plot.background = element_rect(fill = "white", color = NA),
          plot.title = element_text(size = 16, face = "bold"),
          plot.subtitle = element_text(size = 12, face = "italic"),
          plot.caption = element_text(face = "italic"),
          axis.title = element_text(face = "bold"),
          axis.text = element_text(size = 12),
          legend.title = element_text(face = "bold"),
          strip.text = element_text(face = "bold"),
          strip.background = element_rect(fill = "grey80", color = NA))
}
```

# What is data? {#sec-what-is-data}

*Last modified on `r format(fs::file_info("chapter-02-what-is-data.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

## Generale background

## Theoretical background

## R packages used

## Data

```{r}
jump_weight_tbl <- tibble(weight = c(0.6, 1, 2.3, 3.5, 5.2, 7.1, 8.4, 9.2, 10),
                          jump_length = 0.15*weight^3 - 2.2*weight^2 + 8.8*weight + 3.2)
```

@eq-jump-weight

$$
f(x) = 0.15\cdot x^3 - 2.2\cdot x^2 + 8.8 \cdot x + 3.2
$$ {#eq-jump-weight}

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 7
#| fig-cap: "foo. *[Zum Vergrößern anklicken]*"
#| label: fig-02-jump-weight

ggplot(jump_weight_tbl, aes(x = weight, y = jump_length)) +
  theme_book() +
  geom_point()

```


## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Data Science" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}