```{r echo = FALSE}
source("init.R")
```

# What is science? {#sec-what-is-science}

*Last modified on `r format(fs::file_info("chapter-01-what-is-science.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Reality is negotiable." --- Tim Ferriss*



```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-vulcan
#| fig-align: center
#| fig-height: 3
#| fig-width: 7
#| fig-cap: "foo*"

ggplot() +
  theme_void() +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 10.5, b = 4.9, angle = pi/5.5),
               color = col_pal(1, 0.75), fill = col_pal(1, 0.25), linetype = 11) +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 10, b = 5, angle = pi/5.5),
               color = "gray50", fill = "white") +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 5.5, b = 3.5, angle = -2.25),
               color = col_pal(1, 0.75), linetype = 11) +
  annotate("point", x = 0, y = 0, size = 10) +
  annotate("text", x = 0.5, y = 1, hjust = "left", label = "Sun", size = 4) +
  annotate("point", x = 8.6, y = 5, size = 3) +
  annotate("text", x = 8.8, y = 5.5, hjust = "left", label = "Mercury", size = 4) +
  annotate("point", x = 3.5, y = 4.2, size = 3,
           shape = 21, fill = col_pal(1)) +
  annotate("text", x = 3.7, y = 4.7, hjust = "left", label = "Vulcan", size = 4) +
  annotate("label", x = 2.5, y = -7, label = expression(Deviation~"="~frac(34*"°", 3600*"°")%~~%1.2*"%"),
           hjust = "left", fill = col_pal(1, 0.25)) +
  coord_cartesian(xlim = c(-10, 10), ylim = c(-8,6.5))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-habital
#| fig-align: center
#| fig-height: 3
#| fig-width: 7
#| fig-cap: "foo*"

ggplot() +
  # theme_void() +
  geom_ellipse(aes(x0 = -5, y0 = -3, a = 6, b = 5, angle = pi/4),
               color = "gray50") +
  geom_ellipse(aes(x0 = 6, y0 = 4, a = 14, b = 8, angle = pi/4),
               color = col_pal(1, 0.75)) +
  annotate("point", x = 6, y = 4, size = 35) +
  annotate("point", x = c(5.5, 6, 6.3), y = c(3, 2, 1.8), size = c(2, 1.5, 1), 
           shape = 21, color = "white", fill = "white") +
  annotate("text", x = 7.5, y = 7.5, hjust = "left", label = "Sun", size = 4) + 
  annotate("point", x = -5, y = -3, size = 25, color = col_pal(1)) +
  annotate("point", x = 0, y = 0, size = 10, color = "gray25") +
  coord_cartesian(xlim = c(-10, 10), ylim = c(-10, 10))
```

## General background


> *"Any sufficiently advanced technology is indistinguishable from magic." ---  @clarke1968clarke, Clarke's third law*

Full quote

> *"Reality is negotiable. Outside of science and law, all rules can be bent or broken, and it doesn’t require being unethical." --- Tim Ferriss*

Idea of hypotheses

Science is guessing and falsification

@chalmers2013thing [What is this thing called Science?](https://www.pmf.unizg.hr/_download/repository/1._Dodatna_literatura_%28filozofija_znanosti%29%5B3%5D.pdf)

@feynman1966science [What is science](https://www-oc.chemie.uni-regensburg.de/diaz/media/docs/feynman_what-is-science_68.pdf)

@campbell1952science [What is science?](https://www.whatisscience.info/files/what_is_science_rd1.pdf)

@clark2025models [Models Demystified: A Practical Guide from Linear Regression to Deep Learning](https://m-clark.github.io/book-of-models/)

@cox2017statistical [Statistical Thinking for the 21st Century](https://statsthinking21.github.io/statsthinking21-core-site/index.html)

@deutsch2011beginning [The beginning of infinity: Explanations that transform the world](https://www.academia.edu/download/106176299/rtx120400544p.pdf)

[David Deutsch > Quotes](https://www.goodreads.com/author/quotes/103190.David_Deutsch)

## Theoretical background

## R packages used

## Data

## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Data Science" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}