```{r echo = FALSE}
pacman::p_load(tidyverse, patchwork, knitr)
theme_book <- function() {
  theme_minimal() +
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(size = 16, face = "bold"),
          plot.subtitle = element_text(size = 12, face = "italic"),
          plot.caption = element_text(face = "italic"),
          axis.title = element_text(size = 12, face = "bold"),
          axis.text = element_text(size = 11),
          legend.title = element_text(face = "bold"))
}
```

# What is data? {#sec-what-is-data}

*Last modified on `r format(fs::file_info("chapter-02-what-is-data.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"The limits of my language mean the limits of my world." --- Ludwig Wittgenstein*

## General background

## Theoretical background

|   | Symbol |  Application | Englisch |
|------------------------|:----------------------:|--|------------------------|
| *LHS* | $Y / y$ | | **outcome** / response / endpoint / dependent variable |
| *RHS* | $X / x$ |  | **influencer** / influential variable / risk factor / fixed effect / independent variable |
| *RHS* | $Z / z$ |  | **random effect** |
| *RHS* | $X / x$ | Explanation | **explanator** / explanatory variable |
| *RHS* | $X / x$ | Prediction  | **predictor** / predictive variable |
| *RHS* | $X / x$ | Main effect in multiple model | **focal explanator** or **focal predictor** / focal variable |
| $X / x$ is continuous | $c$ | Modelling | **covariate** |
| $X / x$ is categorical | $f_A$ | Modelling | **factor** $A$ with levels $A.1$ to $A.j$ |

: Summary table of terminology used in my statistical modelling. The terms highlighted in **bold** are used here. (*LHS* = left-hand side, *RHS* = right-hand side) {#tbl-names-regression tbl-colwidths="$$14, 8, 28, 50$$"}

A sentence why we use $y$ and not $x$ for mean and other stuff.

## R packages used

## Data

```{r}
jump_weight_tbl <- tibble(x = c(0.6, 1, 2.3, 3.5, 5.2, 7.1, 8.4, 9.2, 10),
                          y = 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2 + rnorm(9, 0, 0.5)) |>
  mutate_all(round, 1) |> 
  rename(weight_mg = x, jump_length_cm = y)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-02-jump-weight
#| tbl-cap: "foo."

jump_weight_tbl|> 
  kable(align = "c", "pipe")
```



@eq-jump-weight

$$
y = 0.15\cdot x^3 - 2.2\cdot x^2 + 8.8 \cdot x + 3.2
$$ {#eq-jump-weight}

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 10
#| fig-cap: "foo. *[Zum Vergrößern anklicken]*"
#| label: fig-02-science-to-model

p1 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Flea weight in [mg]", y = "Jump length in [cm]",
       title = "Science") +
  theme(axis.text = element_blank())

p2 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_point() +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "weight_mg", y = "jump_length_cm",
       title = "Data")

p3 <- ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_function(fun = \(x) 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2,
                size = 1) +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Influencer (X)", y = "Outcome (Y)",
       title = "Model") +
  theme(axis.text = element_blank())

p1 + p2 + p3 +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 5.5
#| fig-cap: "foo. *[Zum Vergrößern anklicken]*"
#| label: fig-02-science-with-model

ggplot(jump_weight_tbl, aes(x = weight_mg, y = jump_length_cm)) +
  theme_book() +
  geom_point(size = 3, shape = 21, fill = "gray75") +
  geom_function(fun = \(x) 0.15*x^3 - 2.2*x^2 + 8.8*x + 3.2) +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  scale_y_continuous(breaks = seq(7, 21, 2), limits = c(7, 22)) +
  labs(x = "Flea weight in [mg]", y = "Jump length in [cm]",
       title = "Science with data and model") 
```



## Alternatives

Further tutorials and R packages on XXX

## Glossary

term

:   what does it mean.

## The meaning of "Data Science" in this chapter.

-   itemize with max. 5-6 words

## Summary

## References {.unnumbered}
