```{r echo = FALSE, warning = FALSE, message = FALSE}
source("init.R")
source("images/part_1/part_1_data.R")
```

# Blood meal of fleas {#sec-temp}

*Last modified on `r format(fs::file_info("chapter-11-covariate-data.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

## General background

## Theoretical background

## R packages used

## Data



### Linear

```{r}
weight_bloodmeal_n14_tbl <- tibble(x = c(1.2, 2.4, 3.5, 3.9, 2.8, 1.8, 3.1, 
                                       5.1, 4.6, 5.5, 6.2, 7.1, 6.5, 7.6),
                                 y = 2 + 1.2*x + rnorm(14, 0, 2)) |>
  mutate_all(round, 1) |> 
  rename(bloodmeal = x, weight = y)
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-weight-bloodmeal-n14
#| tbl-cap: "foo."

weight_bloodmeal_n14_tbl |> 
  kable(align = "c", theme = "striped")
```



```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-01
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "Relationship between blood meal and body weight."

weight_bloodmeal_n14_tbl |> 
  ggplot(aes(x = bloodmeal, y = weight)) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21, fill = "#0D0887FF") +
  labs(x = "Blood meal [ml]", y = "Body weight [mg]") +
  scale_x_continuous(breaks = 0:10) +
  coord_cartesian(xlim = c(0, 8), ylim = c(0, 15))
```



### Non-linear

$$
y = -0.0365*x^4 + 0.9348*x^3 - 7.396*x^2 + 18.876*x + 12.353
$$

```{r}
jump_bloodmeal_n21_tbl <- tibble(x = c(1.2, 2.4, 3.5, 3.9, 2.8, 1.8, 3.1, 
                                       5.1, 4.6, 5.5, 6.2, 7.1, 6.5, 7.6,
                                       8.6, 9.3, 10.5, 9.8, 11.8, 11.1, 10.2),
                                 y = -0.0365*x^4 + 0.9348*x^3 - 7.396*x^2 + 18.876*x + 12.353 + rnorm(21, 0, 3)) |>
  mutate_all(round, 1) |> 
  rename(bloodmeal = x, jumplength = y)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-jump-bloodmeal-n21
#| tbl-cap: "foo."

jump_bloodmeal_n21_tbl |> 
  kable(align = "c", theme = "striped")
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-04
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "Relationship between blood meal and jumping length."

jump_bloodmeal_n21_tbl |> 
  ggplot(aes(x = bloodmeal, y = jumplength)) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21, fill = "#0D0887FF") +
  labs(x = "Blood meal [ml]", y = "Jump length [cm]") +
  scale_x_continuous(breaks = 0:12) +
  coord_cartesian(xlim = c(0, 12), ylim = c(0, NA))
```

### Not independent

```{r}
library(mvtnorm)

sigma <- matrix(c(4,2,2,3), ncol=2)
x <- rmvnorm(n = 21, mean=c(1,2), sigma=sigma) |> 
  as_tibble() |> 
  rename(weight = V1, bodylength = V2)
colMeans(x)
var(x)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-11-cov-02
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"

x |> 
  ggplot(aes(x = weight, y = bodylength, fill = col_pal(1, 0.5))) +
  theme_book() +
  geom_point(show.legend = FALSE, shape = 21) +
  labs(x = "Body weight [mg]", y = "Body length [mm]",
       fill = "Host") +
  scale_fill_viridis(discrete = TRUE, option = "plasma") 
```


## Data availability

The data is available as txt-Files under <https://github.com/jkruppa/biodatascience>.

```{r}
#| echo: false
#| eval: false

weight_bloodmeal_n14_tbl |> 
  mutate_if(is.numeric, round, 2) |> 
  write_delim("data/weight_bloodmeal_n14.txt")

jump_bloodmeal_n21_tbl |> 
  mutate_if(is.numeric, round, 2) |> 
  write_delim("data/jump_bloodmeal_n21.txt")
```


## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}