```{r echo = FALSE, warning = FALSE, message=FALSE}
source("init.R")
```

# Explorative data analysis {#sec-temp}

*Last modified on `r format(fs::file_info("chapter-41-eda.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

## General background

## Theoretical background




```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-dist-test

ggplot()+
  geom_path(data = tibble(x_raw = seq(-3.5, 3.5, 0.01), y_raw = dnorm(x_raw)), 
            aes(y = -(x_raw*1.4-5), x = (y_raw*15-2)),
            linewidth = 1, color = "black", angle = 90) 
```


## R packages used

## Data

```{r}
#| message: false
#| warning: false
jump_dog_tbl <- read_delim("data/jump_fleas.txt") |> 
  select(host, jumplength) |> 
  filter(host == "dog") 
```

```{r}
jump_dog_tbl
```


## Mean


```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 7
#| fig-cap: "foo"
#| label: fig-mean-example

mean_dog_jump <- mean(jump_dog_tbl$jumplength)

jump_dog_plot_tbl <- jump_dog_tbl |>
  mutate(e = jumplength - mean_dog_jump,
         sign = as_factor(ifelse(sign(e) == -1, "\U2012", "+")))

jump_dog_plot_tbl |> 
  ggplot(aes(x = 1:7, y = jumplength)) +
  theme_book() +
  geom_segment(aes(x = 1:7, xend = 1:7, y = jumplength, yend = mean_dog_jump, 
                   color = sign), show.legend = FALSE, size = 1) +
  geom_text(aes(x = 1:7, y = jumplength - e/2, label = sign, color = sign),
            size = 4, position = position_nudge(0.2), fontface = 2, show.legend = FALSE) +
  geom_hline(yintercept = mean_dog_jump) +
  geom_point() +
  scale_x_continuous(breaks = 1:7) +
  scale_y_continuous(breaks = c(25, 30, mean_dog_jump, 35, 40),
                     labels = c("25", "30", round(mean_dog_jump,1), "35", "40")) +
  labs(x = "Dog flea ID", y = "Jump length in [cm]") +
  scale_color_viridis(discrete = TRUE, option = "plasma", end = 0.9) +
  geom_segment(data = filter(jump_dog_plot_tbl, sign == "\U2012"),
    aes(x = seq(4, 5, length.out = 3), xend = seq(4, 5, length.out = 3), 
                   y = (jumplength-4)/2, yend = (mean_dog_jump-4)/2, 
                   color = sign), show.legend = FALSE, size = 1) +
  geom_segment(data = filter(jump_dog_plot_tbl, sign == "+"),
    aes(x = seq(5.5, 6, length.out = 4), xend = seq(5.5, 6, length.out = 4), 
                   y = (jumplength-4)/2, yend = (mean_dog_jump-4)/2, 
                   color = sign), show.legend = FALSE, size = 1) 

```


## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}