```{r echo = FALSE}
source("init.R")
```

# What is science? {#sec-what-is-science}

*Last modified on `r format(fs::file_info("chapter-01-what-is-science.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Reality is negotiable." --- Tim Ferriss*



```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-vulcan
#| fig-align: center
#| fig-height: 3
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_observed <- ggplot() +
  theme_void() +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 10.5, b = 4.9, angle = pi/5.5),
               color = col_pal(1), fill = col_pal(1, 0.25), linetype = 11) +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 10, b = 5, angle = pi/5.5),
               color = "gray50", fill = "white") +
  annotate("point", x = 0, y = 0, size = 10, fill = c6_pal[6], shape = 21) +
  annotate("text", x = 0.6, y = 1, hjust = "left", label = "Sun", size = 5) +
  annotate("point", x = -8.6, y = -5, size = 4) +
  annotate("text", x = -8.3, y = -4.5, hjust = "left", label = "Mercury", size = 5) +
  annotate("label", x = 2.5, y = -7, 
           label = expression(Deviation~"="~frac(34*"°", 3600*"°")%~~%1.2*"%"),
           hjust = "left", fill = col_pal(1, 0.25)) +
  labs(title = "Observed") +
  coord_cartesian(xlim = c(-9, 9), ylim = c(-8,6.5)) +
  theme(plot.title = element_text(size = 16, face = "bold"))

p_model <- ggplot() +
  theme_void() +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 10.5, b = 4.9, angle = pi/5.5),
               color = col_pal(1), fill = col_pal(1, 0.25), linetype = 11) +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 10, b = 5, angle = pi/5.5),
               color = "gray50", fill = "white") +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 5.5, b = 3.5, angle = -2.25),
               color = col_pal(1), linetype = 11) +
  annotate("point", x = 0, y = 0, size = 10, fill = c6_pal[6], shape = 21) +
  annotate("text", x = 0.6, y = 1, hjust = "left", label = "Sun", size = 5) +
  annotate("point", x = -9.1, y = -5, size = 4) +
  annotate("text", x = -8.8, y = -4.5, hjust = "left", label = "Mercury", size = 5) +
  annotate("point", x = 3.5, y = 4.2, size = 3,
           shape = 21, fill = col_pal(1)) +
  annotate("text", x = 3.7, y = 4.7, hjust = "left", label = "Vulcan", size = 5) +
  labs(title = "Model") +
  coord_cartesian(xlim = c(-9, 9), ylim = c(-8,6.5)) +
  theme(plot.title = element_text(size = 16, face = "bold"))

p_observed + p_model +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))

```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-habital
#| fig-align: center
#| fig-height: 3
#| fig-width: 10.5
#| fig-cap: "foo **(A)** foo **(B)** foo"

p_observed <- ggplot() +
  theme_void() +
  geom_ellipse(aes(x0 = 6, y0 = 4, a = 14, b = 8, angle = pi/4),
               color = col_pal(1, 0.75)) +
  annotate("point", x = 6, y = 4, size = 35, fill = c6_pal[6], shape = 21) +
  annotate("point", x = c(5.5, 6, 6.4), y = c(3, 2, 1.8), size = c(2.5, 2, 1.5), 
           shape = 21, color = "black", fill = "black") +
  annotate("text", x = 7.5, y = 7.5, hjust = "left", label = "Sun", size = 5) + 
  annotate("point", x = -5, y = -3, size = 25, color = col_pal(1)) +
  annotate("text", x = -4, y = 0, hjust = "left", label = "Earth", size = 5) +   
  geom_curve(aes(x = -4.5, y = -3, xend = 5.75, yend = 1.5),
             arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
             color = "gray50", size = 1.5, curvature = 0.2,
  ) +
  annotate("label", x = 1.25, y = -3, hjust = "center", label = "Sunspots", size = 5,
           color = "black") + 
  labs(title = "Observed") +
  coord_cartesian(xlim = c(-7, 8), ylim = c(-9, 9)) +
  theme(plot.title = element_text(size = 16, face = "bold"))

p_model <- ggplot() +
  theme_void() +
  geom_ellipse(aes(x0 = -5, y0 = -3, a = 6, b = 5, angle = pi/4),
               color = c6_pal[2]) +
  geom_ellipse(aes(x0 = 6, y0 = 4, a = 14, b = 8, angle = pi/4),
               color = col_pal(1, 0.75)) +
  annotate("point", x = 6, y = 4, size = 35, fill = c6_pal[6], shape = 21) +
  annotate("point", x = c(5.5, 6, 6.4), y = c(3, 2, 1.8), size = c(2.5, 2, 1.5), 
           shape = 21, color = "black", fill = "black") +
  annotate("text", x = 7.5, y = 7.5, hjust = "left", label = "Sun", size = 5) + 
  annotate("point", x = -5, y = -3, size = 25, color = col_pal(1)) +
  annotate("text", x = -4, y = 0, hjust = "left", label = "Earth", size = 5) +   
  geom_ellipse(aes(x0 = c(-5, -4.75), y0 = c(-4, -2), a = c(0.5, 0.75), b = c(0.25, 0.5), 
                   angle = c(-1, -1)),
               fill = "white", color = "white") +
  annotate("point", x = 0, y = 0, size = 10, color = c6_pal[2]) +
  annotate("text", x = 0.5, y = 1.5, hjust = "left", label = "Moon", size = 5) +
  geom_curve(aes(x = -0.25, y = -0.5, xend = -4.5, yend = -3.5),
             arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
             color = "gray50", size = 1.5, curvature = -0.3,
  ) +
  geom_curve(aes(x = 0.25, y = -0.5, xend = 5.75, yend = 1.5),
             arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
             color = "gray50", size = 1.5, curvature = 0.3,
  ) +
  annotate("label", x = 0.1, y = -1.5, hjust = "center", label = "Cloud gaps", size = 5,
           color = "black") + 
  labs(title = "Model") +
  coord_cartesian(xlim = c(-6, 8), ylim = c(-9, 9)) +
  theme(plot.title = element_text(size = 16, face = "bold"))

p_observed + p_model
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-01-hollow
#| fig-align: center
#| fig-height: 3
#| fig-width: 7
#| fig-cap: "foo **(A)** foo **(B)** foo"

ggplot(tibble(x = 1:10, y = 1:10), aes(x,y )) +
  # theme_void() +
  annotate("point", x = 0, y = -7, size = 25, shape = 17) +
  annotate("segment", x = -5, y = -10, xend = 5, yend = 5,
           size = 3, lineend = "round") +
  labs(title = "Observed") +
  coord_cartesian(xlim = c(-10, 10), ylim = c(-10, 10)) +
  theme(plot.title = element_text(size = 16, face = "bold"))
```


## General background


> *"Any sufficiently advanced technology is indistinguishable from magic." ---  @clarke1968clarke, Clarke's third law*

Full quote

> *"Reality is negotiable. Outside of science and law, all rules can be bent or broken, and it doesn’t require being unethical." --- Tim Ferriss*

Idea of hypotheses

Science is guessing and falsification

@chalmers2013thing [What is this thing called Science?](https://www.pmf.unizg.hr/_download/repository/1._Dodatna_literatura_%28filozofija_znanosti%29%5B3%5D.pdf)

@feynman1966science [What is science](https://www-oc.chemie.uni-regensburg.de/diaz/media/docs/feynman_what-is-science_68.pdf)

@campbell1952science [What is science?](https://www.whatisscience.info/files/what_is_science_rd1.pdf)

@clark2025models [Models Demystified: A Practical Guide from Linear Regression to Deep Learning](https://m-clark.github.io/book-of-models/)

@cox2017statistical [Statistical Thinking for the 21st Century](https://statsthinking21.github.io/statsthinking21-core-site/index.html)

@deutsch2011beginning [The beginning of infinity: Explanations that transform the world](https://www.academia.edu/download/106176299/rtx120400544p.pdf)

[David Deutsch > Quotes](https://www.goodreads.com/author/quotes/103190.David_Deutsch)

## Theoretical background

## R packages used

## Data

## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Data Science" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}