```{r echo = FALSE, warning = FALSE, message=FALSE}
source("init.R")
```

# Interlude I - Gummy bears {#sec-temp}

*Last modified on `r format(fs::file_info("interlude-I-gummibaers.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Gummi Bears! Bouncing here and there and everywhere; High adventure that's beyond compare; They are the Gummi Bears." --- Gummi Bears Theme, Adventures of the Gummi Bears*

## General background

## Theoretical background

## R packages used

## Data

```{r}
gummi_tbl <- read_excel("data/gummibears.xlsx")
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-bears
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_bears_tbl <- read_excel("data/gummibears.xlsx") |> 
  select(year:white)

rbind(head(gummi_bears_tbl, n = 4),
      rep("...", times = ncol(gummi_bears_tbl)),
      tail(gummi_bears_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```

```{r}
gummi_count_tbl <- gummi_bears_tbl %>%
  rowwise() %>% 
  mutate(count_bear = sum(c_across(darkred:white)),
         count_color = sum(c_across(darkred:white) > 0)) |> 
  ungroup() |> 
  select(year, count_bear, count_color)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-bears-count
#| tbl-cap: "Extract of the Gummi Bears dataset."

rbind(head(gummi_count_tbl, n = 4),
      rep("...", times = ncol(gummi_count_tbl)),
      tail(gummi_count_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-sudent
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_student_tbl <- read_excel("data/gummibears.xlsx") |> 
  select(year, most_liked:semester) 

rbind(head(gummi_student_tbl, n = 4),
      rep("...", times = ncol(gummi_student_tbl)),
      tail(gummi_student_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```

## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}