```{r echo = FALSE, warning = FALSE, message=FALSE}
source("init.R")
pacman::p_load(ggridges, ggbeeswarm)
```

# Gummy bears {#sec-temp}

*Last modified on `r format(fs::file_info("interlude-I-gummibaers.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Gummi Bears! Bouncing here and there and everywhere; High adventure that's beyond compare; They are the Gummi Bears." --- Gummi Bears Theme, Adventures of the Gummi Bears*

## General background

## Theoretical background

## R packages used

## Data

```{r}
gummi_tbl <- read_excel("data/gummibears.xlsx") |> 
  mutate(gender = factor(gender, levels = c("w", "m"), labels = c("female", "male"))) 
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-bears
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_bears_tbl <- gummi_tbl |> 
  select(year:count_color) |> 
  rename(color_obs = count_color,
         bear_obs = count_bears)

rbind(head(gummi_bears_tbl, n = 4),
      rep("...", times = ncol(gummi_bears_tbl)),
      tail(gummi_bears_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```

```{r}
#| echo: false
gummi_count_tbl <- gummi_bears_tbl %>%
  rowwise() %>% 
  mutate(bear_exp = sum(c_across(darkred:white)),
         color_exp = sum(c_across(darkred:white) > 0),
         bear_diff = bear_obs - bear_exp,
         color_diff = color_obs - color_exp) |> 
  ungroup() |> 
  select(year, bear_obs, bear_exp, bear_diff, color_obs, color_exp, color_diff)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-color-exp
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_count_tbl |> 
janitor::tabyl(color_exp) |> 
  kable(align = "c", "pipe")
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-bear-exp
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_count_tbl |> 
janitor::tabyl(bear_exp) |> 
  kable(align = "c", "pipe")
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-interlude-I-count-bears-year
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: "."

count_tbl <- gummi_count_tbl |> 
  pull(year) |> 
  janitor::tabyl()

gummi_count_tbl |> 
  ggplot(aes(x = bear_exp, y = as_factor(year),
             fill = as_factor(year))) +
  theme_book() +
  stat_density_ridges() +
  scale_x_continuous(breaks = seq(6, 15, by = 1), limits = c(6, 14),
                     sec.axis = sec_axis(~., breaks = seq(6, 15, by = 1))) +
  labs(x = "Number of bears in a bag", y = "Year") +
  theme(legend.position = "none") +
  scale_fill_viridis(option = "plasma", discrete = TRUE) +
  annotate("label", x = 14, y = c(1, 2, 3, 4, 5, 6, 7, 8), 
           label = str_c("n = ", rev(count_tbl$n))) +
  theme(panel.grid.minor.x = element_blank()) +
  coord_flip()
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-interlude-I-diff
#| fig-align: center
#| fig-height: 3.25
#| fig-width: 7
#| fig-cap: "."

gummi_count_tbl |> 
  select(bear_diff, color_diff) |> 
  stack() |> 
  mutate(ind = factor(ind, labels = c("Difference in bears", "Difference in color"))) |> 
  ggplot(aes(x = values)) +
  theme_book() +
  facet_wrap(~ ind, scale = "free_x") +
  geom_bar() +
  geom_label(stat='count', aes(label=..count..), size = 3) +
  scale_x_continuous(breaks = seq(-5, 5,1)) +
  labs(x = "Difference observed and expected", y = "Count")
```



```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-bears-count
#| tbl-cap: "Extract of the Gummi Bears dataset."

rbind(head(gummi_count_tbl, n = 4),
      rep("...", times = ncol(gummi_count_tbl)),
      tail(gummi_count_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-sudent
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_student_tbl <- gummi_tbl |> 
  select(most_liked, gender, height) 

rbind(head(gummi_student_tbl, n = 4),
      rep("...", times = ncol(gummi_student_tbl)),
      tail(gummi_student_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-interlude-I-hist-height
#| fig-align: center
#| fig-height: 4
#| fig-width: 8.5
#| fig-cap: "."

mean_tbl <- gummi_student_tbl |> 
  group_by(gender) |> 
  summarise(mean = round(mean(height, na.rm = TRUE), 1)) |> 
  na.omit()

p1 <- gummi_student_tbl |> 
ggplot(aes(x = height, fill = gender)) +
  theme_minimal() +
  geom_histogram(position = "dodge", color = "black") +
  scale_fill_viridis(option = "plasma", discrete = TRUE, end = 0.7, direction = -1) +
  labs(x = "", y = "Anzahl", fill = "Geschlecht") +
  scale_x_continuous(breaks = seq(150, 215, by = 5)) +
  scale_y_continuous(breaks = seq(0, 55, 5), limits = c(0, 55)) +
  theme(legend.position = "none")

p2 <- gummi_student_tbl |> 
ggplot(aes(x = height, fill = gender)) +
  theme_minimal() +
  geom_density(alpha = 0.75) +
  labs(x = "", y = "", fill = "Geschlecht") +
  scale_x_continuous(breaks = seq(145, 210, by = 5), limits = c(145, 210)) +
  scale_y_continuous(breaks = seq(0, 0.055, 0.01), limits = c(0, 0.057)) +
  scale_fill_viridis(option = "plasma", discrete = TRUE, end = 0.7, direction = -1) +
  geom_vline(xintercept = mean_tbl$mean, color = "black",
             size = 1) +
  annotate("label", x = mean_tbl$mean, y = 0.057, label = mean_tbl$mean) +
  theme(legend.position = "none") #c(0.85, 0.8))

p1 + p2 +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A') +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-interlude-I-beeswarm-height
#| fig-align: center
#| fig-height: 3.5
#| fig-width: 6
#| fig-cap: "."

gummi_student_tbl |> 
  na.omit() |> 
  ggplot(aes(x = gender, y = height, color = gender)) +
  geom_beeswarm(size = 1) +
  theme_minimal() +
  scale_color_viridis(option = "plasma", discrete = TRUE, end = 0.7, direction = -1) +
  labs(x = "Geschlecht", y = "Körpergröße in [cm]") +
  theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(140, 214, 5))
```
## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}