```{r echo = FALSE, warning = FALSE, message=FALSE}
source("init.R")
pacman::p_load(ggridges)
```

# Gummy bears {#sec-temp}

*Last modified on `r format(fs::file_info("interlude-I-gummibaers.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"Gummi Bears! Bouncing here and there and everywhere; High adventure that's beyond compare; They are the Gummi Bears." --- Gummi Bears Theme, Adventures of the Gummi Bears*

## General background

## Theoretical background

## R packages used

## Data

```{r}
gummi_tbl <- read_excel("data/gummibears.xlsx")
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-bears
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_bears_tbl <- read_excel("data/gummibears.xlsx") |> 
  select(year:count_color) |> 
  rename(color_obs = count_color,
         bear_obs = count_bears)

rbind(head(gummi_bears_tbl, n = 4),
      rep("...", times = ncol(gummi_bears_tbl)),
      tail(gummi_bears_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```

```{r}
gummi_count_tbl <- gummi_bears_tbl %>%
  rowwise() %>% 
  mutate(bear_exp = sum(c_across(darkred:white)),
         color_exp = sum(c_across(darkred:white) > 0),
         bear_diff = bear_obs - bear_exp,
         color_diff = color_obs - color_exp) |> 
  ungroup() |> 
  select(year, bear_obs, bear_exp, bear_diff, color_obs, color_exp, color_diff)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-color-exp
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_count_tbl |> 
janitor::tabyl(color_exp) |> 
  kable(align = "c", "pipe")
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-bear-exp
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_count_tbl |> 
janitor::tabyl(bear_exp) |> 
  kable(align = "c", "pipe")
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-interlude-I-count-bears-year
#| fig-align: center
#| fig-height: 3.25
#| fig-width: 7
#| fig-cap: "."

count_tbl <- gummi_count_tbl |> 
  pull(year) |> 
  janitor::tabyl()

gummi_count_tbl |> 
  ggplot(aes(x = bear_exp, y = fct_rev(as_factor(year)),
             fill = fct_rev(as_factor(year)))) +
  theme_book() +
  stat_density_ridges() +
  scale_x_continuous(breaks = seq(6, 15, by = 1), limits = c(6, 13)) +
  labs(x = "Number of bears in a bag", y = "Year") +
  theme(legend.position = "none") +
  geom_vline(xintercept = c(7, 8, 9, 10, 11, 12), linetype = 2, alpha = 0.5) +
  scale_fill_okabeito() +
  annotate("label", x = 12.5, y = c(1, 2, 3, 4, 5, 6, 7, 8) + 0.7, 
           label = str_c("n = ", rev(count_tbl$n))) +
  theme(panel.grid.minor.x = element_blank())
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-chap-interlude-I-diff
#| fig-align: center
#| fig-height: 3.25
#| fig-width: 7
#| fig-cap: "."

gummi_count_tbl |> 
  select(bear_diff, color_diff) |> 
  stack() |> 
  mutate(ind = factor(ind, labels = c("Difference in bears", "Difference in color"))) |> 
  ggplot(aes(x = values)) +
  theme_book() +
  facet_wrap(~ ind, scale = "free_x") +
  geom_bar() +
  geom_label(stat='count', aes(label=..count..), size = 3) +
  scale_x_continuous(breaks = seq(-5, 5,1)) +
  labs(x = "Difference observed and expected", y = "Count")
```



```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-bears-count
#| tbl-cap: "Extract of the Gummi Bears dataset."

rbind(head(gummi_count_tbl, n = 4),
      rep("...", times = ncol(gummi_count_tbl)),
      tail(gummi_count_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-gummi-sudent
#| tbl-cap: "Extract of the Gummi Bears dataset."

gummi_student_tbl <- read_excel("data/gummibears.xlsx") |> 
  select(most_liked:height) 

rbind(head(gummi_student_tbl, n = 4),
      rep("...", times = ncol(gummi_student_tbl)),
      tail(gummi_student_tbl, n = 4)) |> 
  kable(align = "c", "pipe")
```

## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}