```{r echo = FALSE, warning = FALSE, message=FALSE}
source("init.R")
```

# Template chapter {#sec-temp}

*Last modified on `r format(fs::file_info("template-chapter.qmd")$modification_time, '%d. %B %Y at %H:%M:%S')`*

> *"A quote." --- Dan Meyer*

## General background

```{r}
#| message: false
#| warning: false
jump_weight_tbl <- read_delim("data/jump_weight.txt") |> 
  select(weight, jumplength) |> 
  sample_n(5)

mean_cat_jump <- mean(jump_weight_tbl$jumplength)

jump_weight_plot_tbl <- jump_weight_tbl |>
  mutate(sx = weight - mean(weight),
         sy = jumplength - mean(jumplength))
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-align: center
#| fig-height: 4.5
#| fig-width: 7.5
#| fig-cap: "foo"
#| label: fig-var-2-example

p1_corr <- jump_weight_plot_tbl |> 
  ggplot(aes(x = weight, y = jumplength)) +
  theme_book() +
  coord_equal() +
  geom_point() +
  labs(x = "Cat flea weight [mg]", y = "Jump length in [cm]") +
  geom_tile(aes(x = weight - sx/2, 
                y = jumplength - sx/2, 
                width = sx, height = sx), fill = "#FCA63680",
            alpha = 0.2, color = "#FCA636FF", linewidth = 0.5) 

p2_corr <- jump_weight_plot_tbl |> 
  ggplot(aes(x = weight, y = jumplength)) +
  theme_book() +
  coord_equal() +
  geom_point() +
  labs(x = "Cat flea weight [mg]", y = "Jump length in [cm]") +
  geom_tile(aes(x = weight - sy/2, 
                y = jumplength - sy/2, 
                width = sy, height = sy), fill = "#B12A9080",
            alpha = 0.2, color = "#B12A90FF", linewidth = 0.5)

p1_corr + p2_corr + 
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A', tag_prefix = '(', tag_suffix = ')') &
  theme(plot.tag = element_text(size = 16, face = "bold"))

jump_weight_plot_tbl |> 
  ggplot(aes(x = weight, y = jumplength)) +
  theme_book() +
  coord_equal() +
  geom_point() +
  labs(x = "Cat flea weight [mg]", y = "Jump length in [cm]") +
  geom_tile(aes(x = weight - sx/2, 
                y = jumplength - sy/2, 
                width = sx, height = sy), fill = "#B12A9080",
            alpha = 0.2, color = "#B12A90FF", linewidth = 0.5)

```




## Theoretical background

## R packages used

## Data

## Alternatives 

Further tutorials and R packages on XXX

## Glossary

term

: what does it mean.

## The meaning of "Models of Reality" in this chapter.

- itemize with max. 5-6 words

## Summary

## References {.unnumbered}